---
title: "Map sed traps"
author: "Denise Ong"
date: "2023-07-17"
output: html_document
---

Map for sediment traps. to show the path of the incubation.
Using 2 excel sheets. SalpPOOP - SedTrap - 04-29-20 copy for the start and end date of the sediment trap deployment.
InSitu_Array_beacons copy for the location of the ship at each day

I think station 316 location is wrong. I am replacing with Long 180.187666666667 from excel All_CTDs_bottle data.

```{r setup, include=FALSE}
  knitr::opts_chunk$set(message=FALSE,
                        warning=FALSE,
                        cache=TRUE,
                        tidy=FALSE)
```

```{r}
library(here)
library(ncdf4) # package for netcdf manipulation
library(raster) # package for raster manipulation
# library(rgdal) # package for geospatial analysis
library(ggplot2) # package for plotting
library(lattice)
library(RColorBrewer)
library(dplyr)
library(tidyverse)
library(broom)
library(lubridate)
library(ggOceanMaps)
library(tibbletime)
library(ggspatial)
library(phyloseq)
```

#map with stations

## Coordinates of stations
```{r}
location <- readxl::read_excel(here("data", "InSitu_Array_beacons copy.xlsx"), sheet = "data+plots") %>%
  rename(station = 'Stn #',
         cycle = 'Cycle #',
         lat = Latitude,
         long = Longitude) %>%
  mutate(date = date(DeviceDateTime)) %>%
  mutate(time = hour(DeviceDateTime)) %>%
  mutate(datetime = as_datetime(DeviceDateTime)) %>%
  mutate(cycle_name = recode(cycle, "1" = "SA-Sc",
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2"))

# subset to only the dates that the sediment traps are deployed - show the path of the sed traps deployment
SA1_a <- location %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-10-24 16' ~ '2018-10-28 5') %>%
  mutate(sed_traps = "yes")

SA2 <- location %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-11-02 5' ~ '2018-11-06 7') %>%
  mutate(sed_traps = "yes")

ST1 <- location %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-11-07 9' ~ '2018-11-10 7')%>%
  mutate(sed_traps = "yes")


ST2 <- location %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-11-12 7' ~ '2018-11-15 6')%>%
  mutate(sed_traps = "yes")

SA3 <- location %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-11-16 6' ~ '2018-11-19 5')%>%
  mutate(sed_traps = "yes")


sampling_location <- list(SA1_a, SA2, SA3, ST1, ST2) 
sampling_location <- Reduce(
  function(x, y, ...) merge(x, y, all = TRUE, ...),
  sampling_location
)
sampling_location <- sampling_location %>%
  mutate(sample = "sed_traps")


#location of the rest of the sample types
source(here("init_files", "read_data_v2.R")) #load phyloseq and modify variables

location_others <- data.frame(ps_all@sam_data)
location_others <- location_others %>%
  filter(sample %in% c("core", "water_CTD", "water_pump")) %>%
  dplyr::select(latitude, longitude, sample, cycle_name_2, station_id) %>%
  distinct() %>%
  mutate(sed_traps = "no") %>%
  rename(station = station_id,
         cycle_name = cycle_name_2,
         lat = latitude,
         long = longitude) %>%
  mutate(station = as.double(station)) %>%
  mutate(long = case_when(station == 316 ~ 180.1876666, # I think the location listed is wrong, changing based on excel All_CTDs_bottle data
                          TRUE ~ as.numeric(long)))

location_all <- full_join(sampling_location, location_others) %>%
  filter(sed_traps == "yes") # only sediment traps 
```

# map coordinates
```{r}

p <- basemap(limits = c(172, -179.5, -46, -42), rotate = TRUE,  
             lon.interval = 2, lat.interval = 1, grid.size = 0.5, 
             bathymetry = TRUE, bathy.style = "rcg", 
             land.col = "#eeeac4") + 
  geom_spatial_point(data = location_all, aes(x = long, y = lat, shape = sed_traps, size = sed_traps)) +
  scale_shape_manual(values = c(19)) +
  scale_size_manual(values = c(0.2))

p


p <- basemap(limits = c(172, -179.5, -46, -42), rotate = TRUE,  
             lon.interval = 2, lat.interval = 1, grid.size = 0.5, 
             bathymetry = TRUE, bathy.style = "rcg", 
             land.col = "#eeeac4") + 
  geom_spatial_point(data = sampling_location, aes(x = long, y = lat), size = 0.2, shape = 19)

p


# p <- basemap(limits = c(172, -179.5, -46, -42), rotate = TRUE,  
#              lon.interval = 2, lat.interval = 1, grid.size = 0.5, 
#              bathymetry = TRUE, bathy.style = "rcg", 
#              land.col = "#eeeac4") + 
#   geom_spatial_point(data = location_all, aes(x = long, y = lat, shape = sed_traps, size = sed_traps)) +
#   scale_shape_manual(values = c(10, 19)) +
#   scale_size_manual(values = c(1, 0.2))
# 
# p

 pdf("sed traps_map_bathy6.pdf", height=5, width=7) ; plot(p) ; dev.off()
 
#ignore - the updated version of ggoceanmaps already has higher quality bathy map
 # to get better resolution for bathemetry file
#  rb <- raster_bathymetry(bathy = "data/ETOPO1_Ice_g_gmt4.grd",
#                         depths = c(50, 100, 200, 300, 500, 1000, 1500, 2000, 3000, 4000, 5000, 6000), 
#                         proj.out = "EPSG:32636", 
#                         boundary =  c(172, -179, -46, -41.5)
# )
# 
# class(rb)
# names(rb)
# raster::plot(rb$raster)
# 
# bs_bathy <- vector_bathymetry(rb, smooth = TRUE)
# sp::plot(bs_bathy)
```
