---
title: "Table_env"
author: "Denise Ong"
date: "2023-07-23"
output: html_document
---
For this table, I am including the following variables:
- sea surface temp
- surface salinity
- Surface NO3
- Surface integrated PP
- surface integrated Chla

separating SA1-A and SA1-B, because fixed traps and sediment are collected in A and pump is in B. 

# Initialize
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(here)

source(here("init_files", "init.R")) #for libraries
# source(here("init_files", "init_markdown.R")) #for markdown libraries
```



# CTD - salinity and temp

- surface salinity and temp
- only CTD sampled between 9am -12 pm
- cycle 1/subantarctic 1 - 24 Oct to 30 Oct
- cycle 2/subantarctic 2 - 2 Nov to 6 Nov
- cycle 3/subtropical 1 - 8 to 10 Nov
- cycle 4/ subtropical 2 - 12 to 14 Nov
- cycle 5/ subantarctic 3 - 16 to 17 Nov

##Read table
```{r}
CTD_raw <- readxl::read_excel("data/TK_CTD Downcasts copy.xlsx") %>%
  dplyr::rename(CTD = `cast`) %>%
  mutate(datetime =  dmy_hms(datetime)) %>%
  filter(!is.na(t)) %>%
  mutate(date = date(datetime)) %>%
  mutate(time = hour(datetime)) %>%
  mutate(datetime = as_datetime(datetime)) %>%
  mutate(julian_day = julian(date, origin = date("2018-01-01"))) %>%
  filter(time %in% (9:12) ) %>% # only morning CTD
  mutate(t_mean = (t+t2)/2) %>% # each CTD has 2 readings. use the average of the two readings.
  mutate(s_mean = (s+s2)/2) %>%
  # mutate(os_mean = (os+os2)/2) %>%
  select(-t, -t2, -s, -s2, -os, -os2) %>%
  filter(p <= 10) 
CTD_raw

#check CTD,  date, time
check <- readxl::read_excel("data/TK_CTD Downcasts copy.xlsx") %>%
  dplyr::rename(CTD = `cast`) %>%
  mutate(datetime =  dmy_hms(datetime)) %>%
  filter(!is.na(t)) %>%
  mutate(date = date(datetime)) %>%
  mutate(time = hour(datetime)) %>%
  select(date,time) %>%
  distinct()
```

```{r}


# subset to only the dates of sampling


SA1_A <- CTD_raw %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-10-24 16' ~ '2018-10-28 5') %>%
  # filter_time('2018-10-24 16' ~ '2018-10-30 5')  %>% # changing to the whole SA1 duration to include CTD water sample
  mutate(cycle_name = "SA1-a")

SA1_B <- CTD_raw %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-10-28 9' ~ '2018-10-30 5') %>%
  # filter_time('2018-10-24 16' ~ '2018-10-30 5')  %>% # changing to the whole SA1 duration to include CTD water sample
  mutate(cycle_name = "SA1-b")

SA2 <- CTD_raw %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-11-02 5' ~ '2018-11-06 7') %>%
  mutate(cycle_name = "SA2")

ST1 <- CTD_raw %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-11-07 9' ~ '2018-11-10 7') %>%
  mutate(cycle_name = "ST1")

ST2 <- CTD_raw %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-11-12 7' ~ '2018-11-15 6')  %>%
  mutate(cycle_name = "ST2")

SA3 <- CTD_raw %>%
  as_tbl_time(index = datetime) %>%
  filter_time('2018-11-16 6' ~ '2018-11-19 5') %>%
  mutate(cycle_name = "SA3")


CTD <- list(SA1_A, SA1_B, SA2, SA3, ST1, ST2) 
CTD <- Reduce(
  function(x, y, ...) merge(x, y, all = TRUE, ...),
  CTD
)

#check the dates
check <- CTD  %>% 
  select(date, cycle_name, time, CTD, julian_day) %>%
  distinct()
check
```


#Nutrients
## Read table 
combine with C14 hot for the cycle, exp and station number.
```{r}
nutrients_raw <- readxl::read_excel("data/TAN1810_nutrients copy.xls") %>%
  dplyr::rename(CTD = `U_Cast`,
                depth =`Depth`)

CTD_raw <- readxl::read_excel("data/TK_CTD Downcasts copy.xlsx") %>%
  dplyr::rename(CTD = `cast`) %>%
  mutate(datetime =  dmy_hms(datetime)) %>%
  filter(!is.na(t)) %>%
  mutate(date = date(datetime)) %>%
  mutate(time = hour(datetime)) %>%
  mutate(julian_day = julian(date, origin = date("2018-01-01"))) %>%
  select(CTD, datetime, date, time, julian_day) %>%
  mutate(CTD=as.character(CTD)) %>%
  distinct()


nutrients_raw<- full_join(nutrients_raw, CTD_raw) %>%
  filter(!is.na(depth)) %>%
  mutate(date = case_when(CTD == 9168 ~ "2018-11-18", # adding only details for this CTD because want to use, the rest will remove
                          TRUE ~ as.character(date))) %>%
  mutate(date = date(date)) %>%
  mutate(time =  case_when(CTD == 9168 ~ 10, 
                          TRUE ~ as.integer(time))) %>%
  filter(!is.na(date)) %>%
  filter(time %in% (10:12))


```

```{r}


# subset to only the dates of sampling
SA1_A <- nutrients_raw %>%
  as_tbl_time(index = date) %>%
  filter_time('2018-10-24' ~ '2018-10-28') %>%
  # filter_time('2018-10-24 16' ~ '2018-10-30 5')  %>% # changing to the whole SA1 duration to include CTD water sample
  mutate(cycle_name = "SA1-a")

SA1_B <- nutrients_raw %>%
  as_tbl_time(index = date) %>%
  filter_time('2018-10-28' ~ '2018-10-30') %>%
  # filter_time('2018-10-24 16' ~ '2018-10-30 5')  %>% # changing to the whole SA1 duration to include CTD water sample
  mutate(cycle_name = "SA1-b")

SA2 <- nutrients_raw %>%
  as_tbl_time(index = date) %>%
  filter_time('2018-11-02' ~ '2018-11-06') %>%
  mutate(cycle_name = "SA2")

ST1 <- nutrients_raw %>%
  as_tbl_time(index = date) %>%
  filter_time('2018-11-07' ~ '2018-11-10') %>%
  mutate(cycle_name = "ST1")

ST2 <- nutrients_raw %>%
  as_tbl_time(index = date) %>%
  filter_time('2018-11-12' ~ '2018-11-14')  %>%
  mutate(cycle_name = "ST2") %>%
  filter(!(date == "2018-11-15"))

SA3 <- nutrients_raw %>%
  as_tbl_time(index = date) %>%
  filter_time('2018-11-16' ~ '2018-11-19') %>%
  mutate(cycle_name = "SA3")


nutrients_raw <- list(SA1_A, SA1_B, SA2, SA3, ST1, ST2) 
nutrients_raw <- Reduce(
  function(x, y, ...) merge(x, y, all = TRUE, ...),
  nutrients_raw
)

#check the dates
check <- nutrients_raw  %>% 
  select(date, cycle_name, time, CTD) %>%
  distinct()
check

depth_check <- nutrients_raw %>%
  select(cycle_name, depth) %>%
  distinct() %>%
  group_by(cycle_name) %>%
  arrange(cycle_name,depth) 

depth_check

nutrients <- nutrients_raw %>%
  select(CTD, Depth_nominal, depth, NO3_uM, date, time, cycle_name) %>%
  filter(depth <=70)
```

```{r}
summary_sst <-  CTD %>%
  pivot_longer(cols = c("t_mean", "s_mean"), names_to = "ctd", values_to = "value") %>%
  group_by(cycle_name, ctd) %>%
  summarise(mean = round(mean(value), 2),
            sd= round(sd(value),2)) %>%
  unite("mean_sd", mean:sd, sep = " ± ") %>%
  ungroup() %>%
  pivot_wider(names_from = "ctd", values_from = "mean_sd")
summary_sst

summary_nitrate <-   nutrients %>%
  group_by(cycle_name) %>%
  summarise(conc_mean = round(mean(NO3_uM), 2),
            conc_sd= round(sd(NO3_uM),2)) %>%
  unite("NO3_um", conc_mean:conc_sd, sep = " ± ") 
summary_nitrate

summary_all <- full_join(summary_sst, summary_nitrate)
summary_all
```


# NPP

```{r}
npp_raw <- readxl::read_excel("data/TAN1810_NPP_areal_2.1.xlsx", sheet = "Summary_2")  %>%
  select(Cycle, Station, Depth, U_Cast, mean_NPP, sd_NPP, areal_PP) %>%
  distinct() %>%
  mutate(cycle_name = recode(Cycle, "1" = "SA1",
                                    "2" = "SA2",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA3")) %>%
  dplyr::rename(cycle = `Cycle`,
         station = `Station`,
         depth = `Depth`,
         CTD = `U_Cast`) %>%
  mutate(cycle_name = case_when(station == "15" ~ "SA1-a",
                                station == "24" ~ "SA1-a",
                                station == "39" ~ "SA1-a",
                                station == "51" ~ "SA1-b",
                                station == "69" ~ "SA1-b",
                                station == "90" ~ "SA1-b",
                                TRUE ~ as.character(cycle_name)))

# npp_raw$CTD <- sub("^", "U", npp_raw$CTD)

npp_summary <- npp_raw %>%
  select(cycle_name, areal_PP) %>%
  distinct() %>%
  # mutate(water_mass = recode(cycle_name, 
  #                            "SA1" = "SA",
  #                            "SA2" = "SA",
  #                            "SA3" = "SA",
  #                            "ST1" = "ST",
  #                            "ST2" = "ST"
  #                            )) %>%
  # group_by(water_mass) %>%
  group_by(cycle_name) %>%
  summarise(mean = round(mean(areal_PP,  na.rm=TRUE), 1),
            sd = round(sd(areal_PP,  na.rm=TRUE), 1) %>%
   unite("npp_int", mean:sd, sep = " ± ") 
  
npp_summary 
summary_all <- full_join(summary_all, npp_summary)
summary_all
```

# Chla
```{r}
chl <- readxl::read_excel("data/TAN1810_SFChla_2.0 copy.xls", sheet = "chla_int_2") %>%
  dplyr::select(NIWA_ID, U_Cast, cycle, cycle_2, time, depth_corrected, areal_chl)%>%
  filter(!is.na(cycle)) %>%
  dplyr::rename(depth = `depth_corrected`,
         CTD = `U_Cast`) %>%
  mutate(cycle_name = recode(cycle_2, "1a" = "SA1-a",
                                    "1b" = "SA1-b",
                                    "2" = "SA2",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA3")) %>%
  filter(time == "noon")

#by water mass
chl_summary <- chl %>%
  select(cycle_name, areal_chl) %>%
  distinct() %>%
  group_by(cycle_name) %>%
  summarise(mean = round(mean(areal_chl,  na.rm=TRUE), 1),
            sd = round(sd(areal_chl,  na.rm=TRUE), 1) 
            ) %>%
   unite("chl_int", mean:sd, sep = " ± ") 

chl_summary

summary_all <- full_join(summary_all, chl_summary) %>%
  t()

summary_all <- data.frame(summary_all) %>%
  rownames_to_column(var = "var")
summary_all

writexl::write_xlsx(summary_all, path = "env_table_v2.xlsx")
```


