---
title: "diversity analysis"
author: "Denise Ong"
output: html_document
date: "2024-06-27"
---


Using this guide as reference to use microviz: https://david-barnett.github.io/microViz/articles/learn/exercises.html
- Initial checks of the community - trap blanks, replicates and outlier results of fixed traps in SA2.
- Check ASVs with low abundance reads
- subset data options
- Main figure - diversity and richness ASV, overlap between sample types
- supplementary figure - diversity and richness of ASVs between cycles

# Initialize
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(here)

source(here("init_files", "init.R")) #for libraries
source(here("init_files", "init_markdown.R")) #for markdown libraries
source(here("init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps
# cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") #another colour-blind palette, if the previous palatte has colours that are too close together.


#Colours for cycle
colours_cycle <- readxl::read_excel(here("init_files", "colours_cycle.xlsx"))
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)

set.seed(1)
```


# load data
```{r}
source(here("init_files", "read_data_v3.R")) #load phyloseq and modify variables
ps_all
```

## Initial checks 
###  Blanks
blanks: not open during deployment. has a completely different composition.
```{r}
# sample <- as.data.frame(ps_all@sam_data)
plot_list <- ps_all %>%
  ps_filter(sample == "trap_blank_fix" | sample == "trap_fix" | sample == "trap_blank_live" | sample == "trap_live") %>%
  ps_mutate(fixing = case_when(substrate == "sediment trap blank" ~ "blank",
                               substrate == "sediment trap material" ~ "sample")) %>%
  ps_mutate(sample_trap = case_when(sample == "trap_blank_fix" ~ "Fixed traps", 
                                    sample == "trap_fix" ~ "Fixed traps",
                                    sample == "trap_blank_live" ~ "Live traps",
                                    sample == "trap_live" ~ "Live traps")) %>%
  ps_arrange(sample_trap, fixing) %>%
  comp_barplot(
    tax_level = "class", n_taxa = 15,
    label = "fixing",
    group_by = "sample_trap"
  )

# Plot them side by side with the patchwork package.
patch <- patchwork::wrap_plots(plot_list, nrow = 1, guides = "collect")
patch <- patch & facet_grid(fct_relevel(cycle_name_2, "ST1", "ST2", "SA1-A", "SA1-B", "SA2", "SA3") ~ ., space = "free", scales = "free") & coord_flip()
patch <-  patch + plot_annotation(tag_levels = 'A') # make all plots horizontal (note: use & instead of +)
patch
# pdf("checks_blank traps.pdf", height= 10, width=10) ; plot(patch) ; dev.off()

```


```{r}
#fixed †raps
p <- ps_all %>%
   ps_filter(sample == "trap_blank_fix" | sample == "trap_fix") %>% # filter after tax mutate
  ps_arrange(substrate) %>%
  comp_barplot(
    tax_level = "class", n_taxa = 15, merge_other = FALSE,
    label = "substrate_description_detailed"
    ) +
    coord_flip() +
  facet_grid(cycle_name ~., space = "free", scales = "free")
p
# pdf("figs_microviz/checks_blank fixed traps.pdf", height= 10, width=10) ; plot(p) ; dev.off()


# Blanks - live traps
# same, more chrysophyceae in blanks

p <- ps_all %>%
   ps_filter(sample == "trap_blank_live" | sample == "trap_live") %>% # filter after tax mutate
  # ps_arrange(substrate) %>%
  comp_barplot(
    tax_level = "class", n_taxa = 15, merge_other = FALSE,
    label = "substrate_description_detailed"
    ) +
    coord_flip() +
  facet_grid(cycle_name ~., space = "free", scales = "free")
p
# pdf("figs_microviz/checks_blank live traps.pdf", height= 10, width=10) ; plot(p) ; dev.off()

```

### Replicates
replicates look similar. using only one
```{r}
p <- ps_all %>%
  ps_mutate(replicate = case_when(
                                  sample_code == "C4-2-T4METAB0-2" ~ 1,
                                  sample_code == "C4-2-T4METAB0-2TOTAL1-2" ~ 2,
                                  sample_code == "C4-2-T4SFMETAB201-2" ~ 1,
                                  sample_code == "C4-2-T7-Fix-SFDNA20" ~ 2,
                                  sample_code == "C2-1-T9-Live-SFDNA0-2-1-2" ~ NA,
                                  sample_code == "C2-1-T9-Live-TDNA-0-2-1-2" ~ NA,
                                  sample_code == "C5-110-McClane-DNA-1" ~NA,
                                  sample_code == "C5-140-McClane-DNA-2" ~ NA,
                                  TRUE ~ as.numeric(replicate))) %>%
   ps_filter(replicate == 1|replicate == 2) %>% 
  ps_arrange(sample_code) %>%
  comp_barplot(
    tax_level = "class", n_taxa = 15, merge_other = FALSE,
    label = "sample_code"
    ) +
    coord_flip() +
  facet_grid(replicates_check ~., space = "free", scales = "free") +
  ggtitle("Check replicates") +
  theme(strip.text = element_blank())

p

# pdf("checks_replicates.pdf", height= 10, width=8) ; plot(p) ; dev.off()
```

### Fixed traps SA2
Check fixed traps SA2 depth 540 m because very low Shannon diversity for both micro and total. Might be a clogged filter. Removing from samples.
```{r}
p <- ps_all %>%
  ps_filter(sample == "trap_fix", cycle_name == "SA3", depth == "540") %>%
  comp_barplot(
    tax_level = "unique", n_taxa = 15, merge_other = FALSE,
    label = "sample_name"
    ) +
    coord_flip()
p
# pdf("figs_microviz/checks_SA3 fixed 540m.pdf", height= 5, width=10) ; plot(p) ; dev.off()

ps_all %>%
  ps_filter(sample == "trap_fix", cycle_name == "SA3") %>%
  ps_calc_diversity(rank = "unique", index = "shannon") %>%
  samdat_tbl() 

# compare against all of SA3 fixed traps
p <- ps_all %>%
  # subset_taxa(!(ASV == "asv_047_00008")) %>%
  ps_filter(sample == "trap_fix", cycle_name == "SA3") %>%
  comp_barplot(
    tax_level = "unique", n_taxa = 20, merge_other = FALSE,
    label = "sample_name"
    ) +
    coord_flip() +
  facet_grid(depth ~ fraction_name, space = "free", scales = "free")
p
# pdf("figs_microviz/checks_SA3 fixed.pdf", height= 5, width=10) ; plot(p) ; dev.off()

```




## Subset data
remove blank samples, checks and relevel depth
```{r}
# remove blank samples
ps <-  ps_all %>%
  ps_filter(sample != "trap_blank_fix") %>%
  ps_filter(sample != "trap_blank_live")

# remove replicate samples
ps <- ps %>%
  ps_dedupe(method = "random", #set.seed(1) for reproducibility
            vars = c("replicates_check", "station_id"),
            .message_IDs = TRUE) #using dedupe function to randomly select one replicate

# Remove fixed traps SA3 540 m. Dominated by 1 ASV.
ps <-  ps %>%
  ps_filter(sample_name != "C5-4-T4-Fix-TDNA-0-2") %>%
  ps_filter(sample_name != "C5-4-T7-Fix-SFDNA20")

# remove CTD samples that are sampled lower than 40 m
ps <- ps %>%
  ps_filter(!(sample == "water_CTD" & depth >=41))


# order depths
ps <- ps %>% 
  ps_mutate(depth = factor(depth, levels = unique(depth))) %>%
  ps_mutate(depth = fct_relevel(depth, c("428", "780", "1051", "1405", "2604", "540", "500", "340","300", "200", "140", "110", "100", "70", "40", "30", "25", "20", "12", "5"))) %>%
  ps_mutate(depth_category = factor(depth_category, levels = unique (depth_category))) %>%
  ps_mutate(depth_category = fct_relevel(depth_category, c("6", "5", "4", "3", "2", "1")))


# order cycles
ps <- ps %>% 
  ps_mutate(cycle_name = factor(cycle_name, levels = unique(cycle_name))) %>%
  ps_mutate(cycle_name = fct_relevel(cycle_name, c("ST1", "ST2", "SA-Sc-A", "SA-Sc-B","SA1","SA2")))  %>%
  ps_mutate(cycle_name_2 = fct_relevel(cycle_name_2, c("ST1", "ST2", "SA-Sc", "SA1","SA2"))) 

# checks
tax_table(ps) %>% head ()
rank_names(ps)
sample_names(ps) %>% head(10)

# check samples included
check <- data.frame(ps@sam_data) %>%
  select(sample_name, metadata_code, substrate_description_detailed, replicate, replicates_check, date, station_id, fraction_name, depth, cycle_name, cycle_name_2, sample)


ps <- phyloseq_validate(ps)

ps
```


# Sample filtering
low number of reads is from live samples.
```{r}

# total number of reads
ps %>%
  ps_mutate(reads = sample_sums(ps)) %>%
  samdat_tbl() %>%
  summarise(sum = sum(reads))

# reads per sample
ps %>%
  ps_mutate(reads = sample_sums(ps)) %>%
  samdat_tbl() %>%
  ggplot(aes(x = reads)) +
  geom_freqpoly(bins = 500) +
  geom_rug(alpha = 0.5) +
  scale_x_log10(labels = scales::label_number()) +
  labs(x = "Number of classified reads", y = NULL) +
  theme_bw()

```

check low abundance taxa - present in more than 1 sample and more than 100 reads out of 4093870
```{r}
ntaxa(ps) # before filtering 6007 ASVs
ps_filt <- ps %>%
  tax_filter(min_total_abundance = 11) 

ntaxa(ps_filt) # 4572 ASVs
# %>%
#   ntaxa()
ps
ps_filt
```

```{r}
# make table of summary statistics for the unique taxa
TaxaStats <- tibble(
  taxon = taxa_names(ps),
  prevalence = microbiome::prevalence(ps),
  total_abundance = taxa_sums(ps)
)

p <- TaxaStats %>%
  ggplot(aes(total_abundance, prevalence)) +
  geom_point(alpha = 0.5) +
  geom_rug(alpha = 0.1) +
  scale_x_continuous(
    labels = scales::label_number(), name = "Total Abundance"
  ) +
  scale_y_continuous(
    labels = scales::label_percent(), breaks = scales::breaks_pretty(n = 9),
    name = "Prevalence (%)",
    sec.axis = sec_axis(
      trans = ~ . * nsamples(ps), breaks = scales::breaks_pretty(n = 9),
      name = "Prevalence (N samples)"
    )
  ) +
  theme_bw()
p

```

zoom in on less abundant taxa
```{r}
TaxaStats %>%
  ggplot(aes(x = total_abundance, y = prevalence)) +
  geom_vline(xintercept = 10, color = "red", linetype = "dotted") +
  geom_hline(yintercept = 1 / 100, color = "red", linetype = "dotted") +
  geom_point(alpha = 0.5) +
  geom_rug(alpha = 0.1) +
  scale_x_log10(labels = scales::label_number(), name = "Total Abundance") +
  scale_y_log10(
    labels = scales::label_percent(), breaks = scales::breaks_log(n = 9),
    name = "Prevalence (%)",
    sec.axis = sec_axis(
      trans = ~ . * nsamples(ps), breaks = scales::breaks_log(n = 9),
      name = "Prevalence (N samples)"
    )
  ) +
  theme_bw()
```

By taxonomic table
```{r}
ps %>%
  tax_table() %>%
  as.data.frame() %>%
  as_tibble(rownames = "taxon") %>%
  left_join(TaxaStats, by = "taxon") %>%
  add_count(division, name = "division_count", sort = TRUE) %>%
  mutate(division = factor(division, levels = unique(division))) %>% # to fix facet order
  mutate(division = forcats::fct_lump_n(division, n = 5)) %>%
  mutate(division = forcats::fct_explicit_na(division, na_level = "Other")) %>%
  ggplot(aes(total_abundance, prevalence)) +
  geom_vline(xintercept = 10, color = "red", linetype = "dotted") +
  geom_hline(yintercept = 1 / 100, color = "red", linetype = "dotted") +
  geom_point(alpha = 0.5, size = 1) +
  geom_rug(alpha = 0.2) +
  scale_x_log10(
    labels = scales::label_log(), breaks = scales::breaks_log(n = 5),
    name = "Total Abundance"
  ) +
  scale_y_log10(
    labels = scales::label_percent(), breaks = scales::breaks_log(n = 9),
    name = "Prevalence (%)",
    sec.axis = sec_axis(
      trans = ~ . * nsamples(ps), breaks = scales::breaks_log(n = 9),
      name = "Prevalence (N samples)"
    )
  ) +
  facet_wrap("division") +
  theme_bw(10)
```

filter for more than 10 reads per ASV.
```{r}

ps_filt <-  ps %>%
  tax_filter(min_total_abundance = 11) 

ps_filt
check <- as.data.frame(ps_filt@tax_table)
```

# Diversity
From now, working with phyloseq ps, after removing duplicates.


## Figure main - upset plot (ASV presence/absence for sample types)
normalisations and transform to long
```{r}
ps_norm <- phyloseq_normalize_median(ps_filt)
ps_long <- phyloseq_transform_to_long(ps_norm)

#upset only accounts for presence/absence
upset_df <- ps_long %>%
  select(asv_code,sample) %>% # group by the level of classification
  distinct() %>%
  mutate(n=1) %>%
  pivot_wider(names_from = sample,
              values_fill = 0,
              values_from = n) %>%
  data.frame() # this part is important at the end of forming the df if not will have error later

fig_upset_asv <- UpSetR::upset(upset_df, 
                                nsets = 5,
                                empty.intersections = "on", 
                                nintersects = 20,
                                # number.angles = 30, 
                                # point.size = 10, 
                                # line.size = 4, 
                                mainbar.y.label = "ASV number", 
                                sets.x.label = "ASVs", 
                                order.by = "freq", 
                                # text.scale = c(3), 
                                mb.ratio = c(0.6, 0.4))
fig_upset_asv
```

## Figure supp - Shannon diversity and richness (at ASV level)
overall by sample
```{r}
p1<-ps_filt %>%
  ps_calc_diversity(rank = "unique", index = "shannon") %>%
  samdat_tbl() %>%
  mutate(sample = str_replace_all(sample, c(core = "Sediment",
                                        water_CTD = "Water - CTD",
                                        water_pump = "Water - Pump",
                                        trap_live = "Live trap",
                                        trap_fix = "Fix trap"))) %>%
  mutate(sample = fct_relevel(sample, c("Water - CTD", "Water - Pump", "Fix trap", "Live trap", "Sediment"))) %>%
  mutate(depth_type = str_replace_all(sample, c("Sediment" = "Sediment",
                                        "Water - CTD" = "Surface",
                                        "Water - Pump" = "Below DCM (70 m to 500 m)",
                                        "Live trap" = "Below DCM (70 m to 500 m)",
                                        "Fix trap" = "Below DCM (70 m to 500 m)"))) %>%
  mutate(depth_type = fct_relevel(depth_type, c("Surface", "Below DCM (70 m to 500 m)", "Sediment"))) %>%
  ggplot() +
  geom_boxplot(aes(y = shannon_unique, x = sample))+
  # geom_jitter(aes(y = shannon_unique, x = sample, colour = cycle_name_2), width = 0.2, alpha = 0.6) +
  scale_colour_manual(values = colours_cycle) +
  facet_grid(.~ depth_type, scales = "free", space = "free_x") +
  ylab("Shannon diversity") +
  xlab("Sample type") +
  guides(colour=guide_legend(title="Cycle"))+
  theme(legend.position = "top") +
  expand_limits(y = 6)

p1

p2<- ps_filt  %>%
  # tax_filter(min_total_abundance = 11) %>%
  ps_calc_richness(rank = "unique", index = "observed") %>%
  samdat_tbl() %>%
  mutate(sample = str_replace_all(sample, c(core = "Sediment",
                                        water_CTD = "Water - CTD",
                                        water_pump = "Water - Pump",
                                        trap_live = "Live trap",
                                        trap_fix = "Fix trap"))) %>%
  mutate(sample = fct_relevel(sample, c("Water - CTD", "Water - Pump", "Fix trap", "Live trap", "Sediment"))) %>%
  mutate(depth_type = str_replace_all(sample, c("Sediment" = "Sediment",
                                        "Water - CTD" = "Surface",
                                        "Water - Pump" = "Below DCM (70 m to 500 m)",
                                        "Live trap" = "Below DCM (70 m to 500 m)",
                                        "Fix trap" = "Below DCM (70 m to 500 m)"))) %>%
  mutate(depth_type = fct_relevel(depth_type, c("Surface", "Below DCM (70 m to 500 m)", "Sediment"))) %>%
  ggplot() +
  geom_boxplot(aes(y = observed_unique, x = sample))+
  # geom_violin(aes(y = observed_unique, x = sample), outlier.shape = NA) +
  # geom_jitter(aes(y = observed_unique, x = sample, colour = cycle_name_2), width = 0.2, alpha = 0.6) +
  scale_colour_manual(values = colours_cycle) +
  facet_grid(.~ depth_type, scales = "free", space = "free_x") +
  ylab("No. of ASVs") +
  xlab("Sample type") +
  guides(colour=guide_legend(title="Cycle"))+
  theme(legend.position = "none")
p2


p <- p1|p2  + plot_annotation(tag_levels = 'A')
p

# pdf("main_shannon richness sample_boxplot.pdf", height=4, width=8) ; plot(p) ; dev.off()

```



## Figure supp - Shannon diversity and richness by cycle (at ASV level)
```{r}

# barplot group by sample, each bar for cycle. 
p1<-ps_filt %>%
  # tax_filter(min_total_abundance = 11) %>%
  ps_calc_diversity(rank = "unique", index = "shannon") %>%
  samdat_tbl() %>%
  select(sample, cycle_name_2, shannon_unique)  %>%
  group_by(sample, cycle_name_2) %>%
  summarise(mean = mean(shannon_unique),
            sd = sd(shannon_unique)) %>%
  ungroup() %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, c("ST1", "ST2", "SA-Sc", "SA1", "SA2"))) %>%
  mutate(sample = fct_relevel(sample, c("water_CTD", "water_pump", "trap_fix", "trap_live", "core"))) %>%
  ggplot(aes(x = sample, y = mean, fill = cycle_name_2)) +
  geom_bar(stat="identity", color="black", 
           position=position_dodge(), alpha = 0.8) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_manual(values = colours_cycle) +
  ylab("Shannon diversity") +
  xlab("Sample type") +
  guides(fill=guide_legend(title="Cycle"))+
  theme(legend.position = "top")

p1

# barplot group by sample, each bar for cycle. 
p2<-ps_filt %>%
  # tax_filter(min_total_abundance = 11) %>%
  ps_calc_richness(rank = "unique", index = "observed") %>%
  samdat_tbl() %>%
  select(sample, cycle_name_2, observed_unique)  %>%
  group_by(sample, cycle_name_2) %>%
  summarise(mean = mean(observed_unique),
            sd = sd(observed_unique)) %>%
  ungroup() %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, c("ST1", "ST2", "SA-Sc", "SA1", "SA2"))) %>%
  mutate(sample = fct_relevel(sample, c("water_CTD", "water_pump", "trap_fix", "trap_live", "core"))) %>%
  ggplot(aes(x = sample, y = mean, fill = cycle_name_2)) +
  geom_bar(stat="identity", color="black", 
           position=position_dodge(), alpha = 0.8) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_manual(values = colours_cycle) +
  ylab("No. of ASVs") +
  xlab("Sample type") +
  guides(fill=guide_legend(title="Cycle"))+
  theme(legend.position = "none")
p2

p <- p1|p2  + plot_annotation(tag_levels = 'A')
p

# pdf("supp_shannon richness cycle_barplot.pdf", height=4, width=8) ; plot(p) ; dev.off()

```

ignore - some exploratory plots
```{r, eval = FALSE}
# by sample and cycleq
ps_filt  %>%
  ps_calc_diversity(rank = "unique", index = "shannon") %>%
  samdat_tbl() %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, c("ST1", "ST2", "SA-Sc", "SA1", "SA2"))) %>%
  ggplot() +
  geom_boxplot(aes(x = shannon_unique, y = cycle_name_2, colour = cycle_name_2))+
  geom_point(aes(x = shannon_unique, y = cycle_name_2, colour = cycle_name_2), position = position_jitter(height = 0.2), alpha = 0.5) +
  scale_colour_manual(values = colours_cycle)+
  facet_grid(sample~.)

ps_filt  %>%
  ps_calc_diversity(rank = "unique", index = "shannon") %>%
  samdat_tbl() %>%
  ggplot(aes(x = shannon_unique, y = cycle_name, colour = cycle_name)) +
  geom_point(position = position_jitter(height = 0.2), alpha = 0.5) +
  geom_boxplot()+
  scale_colour_manual(values = colours_cycle)


ps_filt %>%
  ps_filter(sample %in% c("trap_fix", "trap_live", "water_pump")) %>%
  ps_calc_diversity(rank = "unique", index = "shannon") %>%
  samdat_tbl() %>%
  ggplot(aes(x = shannon_unique, y = sample, colour = sample)) +
  geom_point(position = position_jitter(height = 0.2), alpha = 0.5) +
  geom_boxplot()+
  # scale_colour_manual(values = colours_cycle) +
  facet_grid(fct_relevel(depth_category, "1", "2", "3", "4")~sample)
```

## statistical tests
at ASV level, live traps have the lowest diversity. CTD and sediment have the highest
fixed traps and water pump have the same diversity.
```{r}
summary_shannon <- ps_filt %>%
  ps_calc_diversity(rank = "unique", index = "shannon") %>%
  samdat_tbl() %>%
  group_by(sample) %>%
  summarise(count = n(),
    mean = mean(shannon_unique, na.rm = TRUE),
    sd = sd(shannon_unique, na.rm = TRUE))
summary_shannon

table <- ps_filt %>%
  ps_calc_diversity(rank = "unique", index = "shannon") %>%
  samdat_tbl()

hist(table$shannon_unique)
qqnorm(table$shannon_unique)
shapiro.test(table$shannon_unique)

# not normally distributed. Use kruskal wallis test

# significantly different diversity between sample type
kruskal.test(shannon_unique~sample, data = table)
library(FSA)  
dunnTest(shannon_unique~sample, data = table,
         method="bonferroni")

# Per cycle
ps_filt %>%
  ps_calc_diversity(rank = "unique", index = "shannon") %>%
  samdat_tbl() %>%
  group_by(sample,cycle_name_2) %>%
  summarise(count = n(),
    mean = mean(shannon_unique, na.rm = TRUE),
    sd = sd(shannon_unique, na.rm = TRUE))

# not significant difference in diversity between cycles
kruskal.test(shannon_unique~cycle_name_2, data = table)
library(FSA)
dunnTest(shannon_unique~cycle_name_2, data = table,
         method="bonferroni")

# library(car)
# lm <- lm(shannon_unique~cycle_name_2, data = table)
# Anova(lm, type = "3")
# pairwise.t.test(table$shannon_unique, table$cycle_name_2, p.adj='bonferroni')

```




# Supp figure - Barplot of all samples

## Custom colour palette - class
```{r}
hueRank <- "division"
hueRankPlural <- "divisions"
shadeRank <- "class"


# Sort phyloseq at lower, and then higher ranks
ps2 <- ps_filt %>%
  tax_sort(by = sum, at = shadeRank) %>%
  tax_sort(by = sum, at = hueRank) %>%
  tax_agg(rank = shadeRank)

# Specify number of hues and shades desired
nHues <- 5 # "Other" phyla will be shades of grey
nShades <- 4 # "Other" families will be the lightest shade of each hue

hierarchicalPalInfo <- data.frame(
  hue = as.vector(tt_get(ps2)[, hueRank]),
  shade = as.vector(tt_get(ps2)[, shadeRank]),
  counts = taxa_sums(otu_get(ps2))
)

hierarchicalPalInfo <- hierarchicalPalInfo %>%
  dplyr::mutate(
    hue = forcats::fct_other(
      f = hue, keep = unique(hue)[seq_len(nHues)],
      other_level = paste("Other", hueRankPlural)
    ),
    nChrHue = nchar(as.character(hue)), padHue = max(nChrHue) - nChrHue
  ) %>%
  dplyr::group_by(hue) %>%
  dplyr::mutate(
    shade = forcats::fct_other(
      f = shade, keep = unique(shade)[seq_len(nShades - 1)],
      other_level = "Other"
    )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    nChrShade = nchar(as.character(shade)), padShade = max(nChrShade) - nChrShade,
    Taxa = paste0(hue, ": ", strrep(" ", padHue), shade, strrep(" ", padShade))
  )
```

```{r}
hierarchicalPalMatrix <- matrix(
  data = sapply(
    X = seq(from = 30, to = 75, length.out = nShades),
    FUN = function(l) scales::hue_pal(l = l, h.start = 30)(n = nHues)
  ),
  byrow = TRUE, ncol = nHues
)
hierarchicalPalMatrix <- cbind(hierarchicalPalMatrix, grey.colors(n = nShades))

hierarchicalPal <- hierarchicalPalMatrix %>%
  as.vector() %>%
  setNames(unique(hierarchicalPalInfo$Taxa))
```
check taxonomic palette
```{r}
tax_palette_plot(hierarchicalPal) +
  theme(axis.text.y.left = element_text(family = "mono"))
```

## core
```{r}
plot_core<- ps2 %>%
  ps_get() %>%
  tax_mutate("division: class" = hierarchicalPalInfo$Taxa, .keep = "none") %>%
   ps_filter(sample == "core") %>% # filter after tax mutate
  comp_barplot(
    tax_level = "division: class", n_taxa = length(hierarchicalPal),
    tax_order = "asis", palette = hierarchicalPal, 
    x = "depth"
    ) +
  facet_grid(fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc-A", "SA-Sc-B", "SA1", "SA2") ~., space = "free", scales = "free") +
  coord_flip() +
  theme(strip.text = element_text(face = "bold")) +
  ggtitle("Sediment core")
plot_core

# pdf("barplot_core.pdf", height=6, width=10) ; plot(plot_core) ; dev.off()
```


## pump
```{r}
plot_pump<- ps2 %>%
  ps_get() %>%
  tax_mutate("division: class" = hierarchicalPalInfo$Taxa, .keep = "none") %>%
   ps_filter(sample == "water_pump") %>% # filter after tax mutate
  comp_barplot(
    tax_level = "division: class", n_taxa = length(hierarchicalPal),
    tax_order = "asis", palette = hierarchicalPal, 
    x = "depth"
    ) +
  facet_grid(fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc-A", "SA-Sc-B", "SA1", "SA2") ~ ., space = "free", scales = "free") +
  coord_flip() +
  theme(
    # axis.ticks.y = element_blank(), 
    strip.text = element_text(face = "bold")) +
  ggtitle("Lower water column (McClane pump)")

plot_pump
# pdf("barplot_pump.pdf", height=6, width=10) ; plot(plot_pump) ; dev.off()

```


##  live
```{r}
plot_live<- ps2 %>%
  ps_get() %>%
  tax_mutate("division: class" = hierarchicalPalInfo$Taxa, .keep = "none") %>%
  ps_filter(sample == "trap_live") %>% # filter after tax mutate
  comp_barplot(
    tax_level = "division: class", n_taxa = length(hierarchicalPal),
    tax_order = "asis", palette = hierarchicalPal, 
    x = "depth"
    ) +
  facet_grid(fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc-A", "SA-Sc-B", "SA1", "SA2") ~ fraction_name
             , space = "free", scales = "free"
             ) +
  coord_flip() +
  theme(
    # axis.ticks.y = element_blank(), 
        strip.text = element_text(face = "bold")) +
  ggtitle("Live traps") +
  xlab(NULL)

plot_live
# pdf("barplot_trap live.pdf", height=6, width=10) ; plot(plot_live) ; dev.off()

```


##  fixed 
```{r}
plot_fix <- ps2 %>%
  ps_get() %>%
  tax_mutate("division: class" = hierarchicalPalInfo$Taxa, .keep = "none") %>%
  ps_filter(sample == "trap_fix") %>% # filter after tax mutate
  comp_barplot(
    tax_level = "division: class", n_taxa = length(hierarchicalPal),
    tax_order = "asis", palette = hierarchicalPal, 
    x = "depth"
    ) +
  facet_grid(fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc-A", "SA-Sc-B", "SA1", "SA2") ~ fraction_name,
             space = "free", scales = "free"
             ) +
  coord_flip() +
  theme(strip.text = element_text(face = "bold")) +
  ggtitle("Fixed traps")

plot_fix
# pdf("barplot_trap fix.pdf", height=6, width=10) ; plot(plot_fix) ; dev.off()

```




##    CTD
```{r}
plot_ctd<- ps2 %>%
  ps_get() %>%
  tax_mutate("division: class" = hierarchicalPalInfo$Taxa, .keep = "none") %>%
   ps_filter(sample == "water_CTD") %>% # filter after tax mutate
  comp_barplot(
    tax_level = "division: class", n_taxa = length(hierarchicalPal),
    tax_order = "asis", palette = hierarchicalPal, 
    x = "depth"
    ) +
  facet_grid(fct_relevel(cycle_station, "ST1_193", "ST1_239", "ST2_266", "ST2_316", "SA-Sc-A_24", "SA-Sc-B_108","SA1-137",  "SA1_188", "SA2-324", "SA2-371") ~ ., space = "free", scales = "free") +
  coord_flip() +
  theme(strip.text = element_text(face = "bold")) +
  ggtitle("Upper water column (CTD)")

plot_ctd
# pdf("barplot_CTD.pdf", height=7, width=10) ; plot(plot_ctd) ; dev.off()

```

```{r}
library(patchwork)
p <- (plot_ctd * theme(legend.position="none") * ylab(NULL) * xlab("Depth") | 
        plot_pump * ylab("Proportion of reads")  * xlab(NULL) | 
        plot_core * theme(legend.position="none") * ylab(NULL) * xlab(NULL)) / 
    (plot_fix * theme(legend.position="none")  * ylab("Proportion of reads") * xlab("Depth")| 
      plot_live * theme(legend.position="none") * ylab(NULL) * xlab(NULL)) + 
  plot_layout(guides = 'collect', heights = c(3, 2)) + plot_annotation(tag_levels = 'A') 

p

# pdf("barplot_all samples_class.pdf", height=14, width=17) ; plot(p) ; dev.off()

```