"ST1", "ST2", "SA-Sc", "SA1", "SA2"---
title: "NMDS"
author: "Denise Ong"
date: "5/22/2023"
output: html_document
---

NMDS output after filtering (Microviz2.Rmd) using rarify to control for uneven sampling. Using PERMANOVA, pairwise PERMANOVA and betadisper to check for significant differerence in community composition.
Note: more separation by sample type instead of sampling location. But some locations are more clustered than others (less community variation)

# Initialize
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("init_files", "init.R")) #for libraries
source(here("init_files", "init_markdown.R")) #for markdown libraries
source(here("init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps 
# source(here("init_files", "colours_random.R")) # for colours, using Rcolourbrewer
# cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") #another colour-blind palette, if the previous palatte has colours that are too close together.

set.seed(1)
```

```{r}
# convert the sample_data() within a phyloseq object to a vegan compatible data object
pssd2veg <- function(physeq) {
  sd <- sample_data(physeq)
  return(as(sd,"data.frame"))
}

# convert the otu_table() within a phyloseq object to a vegan compatible data object
psotu2veg <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}
```

# load data
```{r}
source(here("init_files", "read_data_v3.R")) #load phyloseq and modify variables
ps_all
source(here("init_files", "filt_samples.R")) #filter out replicates, blanks, low read ASVs etc. 
ps
```


# All samples
```{r}
sample <- data.frame(ps@sam_data) %>%
  rownames_to_column(var = "label")

otu <- psotu2veg(ps)


#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)

# subset removed groups and samples from the plot
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$sample_name <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  inner_join(., sample_list)

rownames(otu_sub) <- otu_sub$sample_name

otu_sub <- otu_sub %>%
  select(-sample_name) %>%
  as.matrix()

```


### checks before plotting - ok, ignore for now.
```{r,eval=FALSE}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu_sub %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
    group_by(file_code) %>% # count rads of each sample
    # group_by(otu) %>% # count reads of asvs
    summarise(sum = sum(reads))
otu_long
```

```{r, eval=FALSE}

p <- rarecurve(otu_sub, step = 40, label = FALSE)

pdf("rare curve.pdf", height= 6, width=6) ; plot(p) ; dev.off()

```

## NMDS
```{r}

dist <- avgdist(otu_sub, sample = 1066, dmethod = "bray") #rarify to control for uneven number of reads per sample. lowest number of reads in a sample is 1066.

nmds_sub<- metaMDS(dist)

nmds_sample_sub <- scores(nmds_sub) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

goodness(nmds_sub)
stressplot(nmds_sub)

nmds_sub$stress

#plot by sample type
nmds_sample <- nmds_sample_sub %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA-Sc", "SA1","SA2" )) %>%
    mutate(sample = fct_relevel(sample, "water_CTD", "water_pump", "trap_fix", "trap_live", "core")) %>%
    ggplot() +
    stat_ellipse(aes(x=NMDS1, y= NMDS2, colour = sample)) +
    geom_point(aes(x=NMDS1, y= NMDS2, colour = sample, shape = cycle_name_2
                   ), size =2.5) +
    scale_colour_manual(values = sample_palette) +
    scale_shape_manual(values=c(3,7,16,17,15)) + # shapes for cycle
    theme(legend.position = "right",
        axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=13),
        axis.text.y = element_text(size = 10),
        axis.title.y =element_text(size=13)) +
    annotate("text", x = 0.3, y = -0.5, label = "Stress = 0.20", size = 5)


nmds_sample

pdf("nmds_samples.pdf", height=5, width=6.5) ; plot(nmds_sample) ; dev.off()
```

## PERMANOVA

```{r}
dist_df <- as.data.frame(as.matrix(dist))
dist_df$sample_name <-rownames(dist_df) #row names to column
rownames(dist_df) <- c() #remove row names
meta_dist <-  inner_join(sample, dist_df, by = "sample_name") #join with the metdata
all_dist <- dist_df %>%
    select(all_of(.[["sample_name"]])) %>%
    as.dist() #something to change to a distance matrix for adonis

adonis2(all_dist~sample+water_mass, data=meta_dist)
pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$water_mass)
pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$sample)
# adonis2(all_dist~sample+cycle_name_2, data=meta_dist)
# pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$cycle_name_2)


# betadisper analysis
bd <- betadisper(all_dist, meta_dist$sample)
anova(bd)
permutest(bd)
meta_dist %>% count(sample)

## Tukey's Honest Significant Differences
(mod.HSD <- TukeyHSD(bd))
plot(mod.HSD)

# meta_dist %>% group_by(sample) %>% count(sample)
```

# Water CTD
```{r}
sample <- data.frame(ps@sam_data) %>%
  filter(sample == "water_CTD")

otu <- psotu2veg(ps)

#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)

# subset removed groups and samples from the plot
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$sample_name <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  inner_join(., sample_list)

rownames(otu_sub) <- otu_sub$sample_name

otu_sub <- otu_sub %>%
  select(-sample_name) %>%
  as.matrix()
```

checks before plotting - ok, ignore for now.
```{r, eval=FALSE}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu_sub %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
    group_by(file_code) %>% # count rads of each sample
    # group_by(otu) %>% # count reads of asvs
    summarise(sum = sum(reads))
otu_long
```

NMDS
```{r}
dist <- avgdist(otu_sub, sample = 12623, dmethod = "bray") #rarify to control for uneven number of reads per sample. 

nmds_sub<- metaMDS(dist)

nmds_sample_sub <- scores(nmds_sub) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

goodness(nmds_sub)
stressplot(nmds_sub)

nmds_sub$stress

#plot by cycle and depth
ctd<- nmds_sample_sub %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA-Sc", "SA1", "SA2" )) %>%
    ggplot() +
    stat_ellipse(aes(x=NMDS1, y= NMDS2, linetype = water_mass)) +
    geom_point(aes(x=NMDS1, y= NMDS2, colour = cycle_name_2)) +
    scale_colour_manual(values = colours_cycle) +
    theme_classic() +
    theme(legend.position = "right")  +
  ggtitle("Upper water column")# +
  # annotate("text", x = 0.5, y = -0.3, label = "Stress = 0.13")

ctd

# pdf("nmds_live traps.pdf", height=6, width=7) ; plot(nmds) ; dev.off()
```

PERMANOVA
```{r}
dist_df <- as.data.frame(as.matrix(dist))
dist_df$sample_name <-rownames(dist_df) #row names to column
rownames(dist_df) <- c() #remove row names
meta_dist <-  inner_join(sample, dist_df, by = "sample_name") #join with the metdata
all_dist <- dist_df %>%
    select(all_of(.[["sample_name"]])) %>%
    as.dist() #something to change to a distance matrix for adonis

# meta_dist %>% group_by(water_mass) %>% count(sample)

# no diff for depth and size fraction
adonis2(all_dist~water_mass, data=meta_dist)
pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$water_mass)

# bd <- betadisper(all_dist, meta_dist$cycle_name_2)
# anova(bd) #p=0.001
# permutest(bd)
# meta_dist %>% count(cycle_name_2)
# 
# ## Tukey's Honest Significant Differences
# (mod.HSD <- TukeyHSD(bd)) #SA3-SA2 p=0.002, SA1-SA2 p=0.007
```


# water pump
```{r}
sample <- data.frame(ps@sam_data) %>%
    filter(sample == "water_pump")

otu <- psotu2veg(ps)


#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)

# subset removed groups and samples from the plot
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$sample_name <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  inner_join(., sample_list)

rownames(otu_sub) <- otu_sub$sample_name

otu_sub <- otu_sub %>%
  select(-sample_name) %>%
  as.matrix()

# rarecurve(otu_sub, step = 40, label = FALSE)

```

checks before plotting - ok, ignore for now.
```{r, eval=FALSE}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu_sub %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
    group_by(file_code) %>% # count rads of each sample
    # group_by(otu) %>% # count reads of asvs
    summarise(sum = sum(reads))
otu_long
```

PERMANOVA
```{r}
dist <- avgdist(otu_sub, sample = 12000, dmethod = "bray") #rarify to control for uneven number of reads per sample. 

nmds_sub<- metaMDS(dist)

nmds_sample_sub <- scores(nmds_sub) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

goodness(nmds_sub)
stressplot(nmds_sub)

nmds_sub$stress

#plot by cycle and depth
ctd<- nmds_sample_sub %>%
    # mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA1-A", "SA1-B","SA2","SA3" )) %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA1", "SA2","SA3" )) %>%
    ggplot() +
    stat_ellipse(aes(x=NMDS1, y= NMDS2, 
                     colour = water_mass)
                     #,linetype = water_mass)
                 # , level = 0.75
                 ) +
    geom_point(aes(x=NMDS1, y= NMDS2, colour = cycle_name_2
                   # ,
                   # shape = fct_rev(depth_category)
                   # shape = fraction_name
                   )) +
    scale_colour_manual(values = colours_cycle) +
    # scale_shape_manual(values=c(0,1, 15, 17, 19)) + # shapes for cycle
    # scale_shape_manual(values=c(1, 19)) + # shapes for water masss=
    theme_classic() +
    theme(legend.position = "right") 
  # ggtitle("Lower water column (70 to 540 m)") +
  # annotate("text", x = 1.2, y = -0.3, label = "Stress = 0.13")

ctd

# pdf("nmds_pump.pdf", height=6, width=7) ; plot(nmds) ; dev.off()

```

```{r}
dist_df <- as.data.frame(as.matrix(dist))
dist_df$sample_name <-rownames(dist_df) #row names to column
rownames(dist_df) <- c() #remove row names
meta_dist <-  inner_join(sample, dist_df, by = "sample_name") #join with the metdata
all_dist <- dist_df %>%
    select(all_of(.[["sample_name"]])) %>%
    as.dist() #something to change to a distance matrix for adonis

# meta_dist %>% group_by(water_mass) %>% count(sample)

# PERMANOVA based on cycle, not sure if there is too little samples for this.
adonis2(all_dist~water_mass, data=meta_dist) # adjusted pairwise p values are not significant. I think it is because there are too many groups.

pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$water_mass)

# bd <- betadisper(all_dist, meta_dist$cycle_name_2)
# anova(bd)
# permutest(bd)
# meta_dist %>% count(cycle_name_2)

## Tukey's Honest Significant Differences
# (mod.HSD <- TukeyHSD(bd))
# plot(mod.HSD)

# # PERMANOVA based on water mass ST vs SA
# adonis2(all_dist~water_mass*depth_category_2, data=meta_dist)
# 
# pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$depth_category_2)
# 
# bd <- betadisper(all_dist, meta_dist$depth_category)
# anova(bd)
# permutest(bd)
# meta_dist %>% count(water_mass)

```

# Fixed traps
```{r}
sample <- data.frame(ps@sam_data) %>%
  filter(sample == "trap_fix") #%>%
  # filter(fraction_name != "pico-nano") %>%
  # filter(fraction_name != "total")

otu <- psotu2veg(ps)


#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)

# subset removed groups and samples from the plot
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$sample_name <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  inner_join(., sample_list)

rownames(otu_sub) <- otu_sub$sample_name

otu_sub <- otu_sub %>%
  select(-sample_name) %>%
  as.matrix()
```

checks before plotting - ok, ignore for now.
```{r, eval=FALSE}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu_sub %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
    group_by(file_code) %>% # count rads of each sample
    # group_by(otu) %>% # count reads of asvs
    summarise(sum = sum(reads))
otu_long
```

PERMANOVA
```{r}
dist <- avgdist(otu_sub, sample = 7000, dmethod = "bray") #rarify to control for uneven number of reads per sample. 

nmds_sub<- metaMDS(dist)

nmds_sample_sub <- scores(nmds_sub) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

goodness(nmds_sub)
stressplot(nmds_sub)

nmds_sub$stress

#plot by cycle and depth
fix <- nmds_sample_sub %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA-Sc", "SA1","SA2" )) %>%
    ggplot() +
    geom_point(aes(x=NMDS1, y= NMDS2, 
                   colour = cycle_name_2,
                   shape = fraction_name
                   )) +
  stat_ellipse(aes(x=NMDS1, y= NMDS2, 
                     colour = water_mass)
                     #,linetype = water_mass)
                 # , level = 0.75
                 ) +
    scale_colour_manual(values = colours_cycle) +
    # scale_shape_manual(values=c(19, 0, 2)) + #shapes for size fraction
    theme_classic() +
    theme(legend.position = "right")  +
  # annotate("text", x = 0.5, y = -0.3, label = "Stress = 0.21") +
  ggtitle("Fixed traps")

fix
```

```{r}
dist_df <- as.data.frame(as.matrix(dist))
dist_df$sample_name <-rownames(dist_df) #row names to column
rownames(dist_df) <- c() #remove row names
meta_dist <-  inner_join(sample, dist_df, by = "sample_name") #join with the metdata
all_dist <- dist_df %>%
    select(all_of(.[["sample_name"]])) %>%
    as.dist() #something to change to a distance matrix for adonis

# meta_dist %>% group_by(fraction_name) %>% count(sample)

adonis2(all_dist~water_mass, data=meta_dist) #p=0.043
pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$water_mass)

# bd <- betadisper(all_dist, meta_dist$fraction_name)
# anova(bd)
# permutest(bd)
# meta_dist %>% count(fraction_name)
# 
# ## Tukey's Honest Significant Differences
# (mod.HSD <- TukeyHSD(bd))
# plot(mod.HSD)

```
# Live traps
```{r}
sample <- data.frame(ps@sam_data) %>%
  filter(sample == "trap_live") #%>%
  # filter(fraction_name != "pico-nano") %>%
  # filter(fraction_name != "total")

otu <- psotu2veg(ps)


#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)

# subset removed groups and samples from the plot
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$sample_name <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  inner_join(., sample_list)

rownames(otu_sub) <- otu_sub$sample_name

otu_sub <- otu_sub %>%
  select(-sample_name) %>%
  as.matrix()
```

checks before plotting - ok, ignore for now.
```{r, eval=FALSE}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu_sub %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
    group_by(file_code) %>% # count rads of each sample
    # group_by(otu) %>% # count reads of asvs
    summarise(sum = sum(reads))
otu_long
```

PERMANOVA
```{r}
dist <- avgdist(otu_sub, sample = 1400, dmethod = "bray") #rarify to control for uneven number of reads per sample. 

dist_df <- as.data.frame(as.matrix(dist))
dist_df$sample_name <-rownames(dist_df) #row names to column
rownames(dist_df) <- c() #remove row names
meta_dist <-  inner_join(sample, dist_df, by = "sample_name") #join with the metdata
all_dist <- dist_df %>%
    select(all_of(.[["sample_name"]])) %>%
    as.dist() #something to change to a distance matrix for adonis

# meta_dist %>% group_by(fraction_name) %>% count(sample)

adonis2(all_dist~water_mass, data=meta_dist) #p=0.001
```
# Sediments
```{r}
sample <- data.frame(ps@sam_data) %>%
  filter(sample == "core") 

otu <- psotu2veg(ps)


#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)

# subset removed groups and samples from the plot
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$sample_name <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  inner_join(., sample_list)

rownames(otu_sub) <- otu_sub$sample_name

otu_sub <- otu_sub %>%
  select(-sample_name) %>%
  as.matrix()
```

checks before plotting - ok, ignore for now.
```{r, eval=FALSE}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu_sub %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
    group_by(file_code) %>% # count rads of each sample
    # group_by(otu) %>% # count reads of asvs
    summarise(sum = sum(reads))
otu_long
```

PERMANOVA
```{r}
dist <- avgdist(otu_sub, sample = 11102, dmethod = "bray") #rarify to control for uneven number of reads per sample. 

dist_df <- as.data.frame(as.matrix(dist))
dist_df$sample_name <-rownames(dist_df) #row names to column
rownames(dist_df) <- c() #remove row names
meta_dist <-  inner_join(sample, dist_df, by = "sample_name") #join with the metdata
all_dist <- dist_df %>%
    select(all_of(.[["sample_name"]])) %>%
    as.dist() #something to change to a distance matrix for adonis

# meta_dist %>% group_by(fraction_name) %>% count(sample)

adonis2(all_dist~water_mass, data=meta_dist) #p=0.001
```
# For presentation - exclude live traps
```{r}
sample <- data.frame(ps@sam_data) %>%
  filter(sample != "trap_live") #%>%
  # filter(sample != "core")

otu <- psotu2veg(ps)


#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)

# subset removed groups and samples from the plot
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$sample_name <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  inner_join(., sample_list)

rownames(otu_sub) <- otu_sub$sample_name

otu_sub <- otu_sub %>%
  select(-sample_name) %>%
  as.matrix()

# rarecurve(otu_sub, step = 40, label = FALSE)

```

checks before plotting - ok, ignore for now.
```{r, eval=FALSE}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu_sub %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
    group_by(file_code) %>% # count rads of each sample
    # group_by(otu) %>% # count reads of asvs
    summarise(sum = sum(reads))
otu_long
```

PERMANOVA
```{r}
dist <- avgdist(otu_sub, sample = 7218, dmethod = "bray") #rarify to control for uneven number of reads per sample. 

nmds_sub<- metaMDS(dist)

nmds_sample_sub <- scores(nmds_sub) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

goodness(nmds_sub)
stressplot(nmds_sub)

nmds_sub$stress


#plot by sample type
nmds_sample <- nmds_sample_sub %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA-Sc", "SA1","SA2" )) %>%
    mutate(sample = fct_relevel(sample, "water_CTD", "water_pump", "trap_fix", "trap_live", "core")) %>%
    ggplot() +
    stat_ellipse(aes(x=NMDS1, y= NMDS2, colour = sample)) +
    geom_point(aes(x=NMDS1, y= NMDS2, colour = sample, shape = cycle_name_2
                   )) +
    scale_colour_manual(values = sample_palette) +
    scale_shape_manual(values=c(3,7,16,17,15)) + # shapes for cycle
    theme(legend.position = "right") +
    annotate("text", x = 0.3, y = -0.5, label = "Stress = 0.17")


nmds_sample

pdf("presentation_nmds.pdf", height=6, width=7) ; plot(nmds_sample) ; dev.off()

```

```{r}
dist_df <- as.data.frame(as.matrix(dist))
dist_df$sample_name <-rownames(dist_df) #row names to column
rownames(dist_df) <- c() #remove row names
meta_dist <-  inner_join(sample, dist_df, by = "sample_name") #join with the metdata
all_dist <- dist_df %>%
    select(all_of(.[["sample_name"]])) %>%
    as.dist() #something to change to a distance matrix for adonis

meta_dist %>%  dplyr::count(sample)

# PERMANOVA based on cycle, not sure if there is too little samples for this.
adonis2(all_dist~sample, data=meta_dist) # adjusted pairwise p values are not significant. I think it is because there are too many groups.

pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$sample)

bd <- betadisper(all_dist, meta_dist$sample)
anova(bd)
permutest(bd)
meta_dist %>% count(sample)

## Tukey's Honest Significant Differences
(mod.HSD <- TukeyHSD(bd))
plot(mod.HSD)

# # PERMANOVA based on water mass ST vs SA
# adonis2(all_dist~water_mass*sample, data=meta_dist)
adonis2(all_dist~water_mass+sample, data=meta_dist)


pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$sample)
# 
bd <- betadisper(all_dist, meta_dist$sample)
anova(bd)
permutest(bd)
# meta_dist %>% count(water_mass)

``` 