---
title: "NMDS"
author: "Denise Ong"
date: "5/22/2023"
output: html_document
---

NMDS output after filtering (Microviz2.Rmd) using rarify to control for uneven sampling. Using PERMANOVA, pairwise PERMANOVA and betadisper to check for significant differerence in community composition.
Note: more separation by sample type instead of sampling location. But some locations are more clustered than others (less community variation)

# Initialize
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("init_files", "init.R")) #for libraries
# source(here("init_files", "init_markdown.R")) #for markdown libraries
source(here("init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps 
# source(here("init_files", "colours_random.R")) # for colours, using Rcolourbrewer
# cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") #another colour-blind palette, if the previous palatte has colours that are too close together.

set.seed(1)
```

```{r}
# convert the sample_data() within a phyloseq object to a vegan compatible data object
pssd2veg <- function(physeq) {
  sd <- sample_data(physeq)
  return(as(sd,"data.frame"))
}

# convert the otu_table() within a phyloseq object to a vegan compatible data object
psotu2veg <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}
```

# load data
```{r}
source(here("init_files", "read_data_v3.R")) #load phyloseq and modify variables
ps_all
source(here("init_files", "filt_samples.R")) #filter out replicates, blanks, low read ASVs etc. 
ps
```


# All samples
```{r}
sample <- data.frame(ps@sam_data) %>%
  rownames_to_column(var = "label")

otu <- psotu2veg(ps)


#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)
```


### checks before plotting - ok, ignore for now.
```{r,eval=FALSE}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
    group_by(file_code) %>% # count rads of each sample
    # group_by(otu) %>% # count reads of asvs
    summarise(sum = sum(reads))
otu_long
writexl::write_xlsx(otu_long, ""
```

## rarefaction curve
```{r, eval=FALSE}

p <- rarecurve(otu, step = 40, label = FALSE)
p
# 
# 
# graphics.off()
# pdf("rare curve_all samples.pdf", width=7, height=5)
# rarecurve(otu, step=40, label=FALSE, 
#           xlab = "Sample size",
#           ylab = "Number of ASVs")
# dev.off()


```


### rarecurve for live traps
```{r}
ps_sub <- ps %>%
  ps_filter(sample == "trap_live")

ps_sub
otu_sub <- psotu2veg(ps_sub)

rarecurve(otu_sub, step = 40, label = FALSE)

# cycles <- c("C2", "C3", "C4", "C5") 
# 
# #subset samples for each cycle
# d_c2 <- otu_sub[grep("C2", rownames(otu_sub)), ]
# d_c3 <- otu_sub[grep("C3", rownames(otu_sub)), ]
# d_c4 <- otu_sub[grep("C4", rownames(otu_sub)), ]
# d_c5 <- otu_sub[grep("C5", rownames(otu_sub)), ]
# 
# 
# # pool the asv counts for each cycle
# pool_c2 <- colSums(d_c2)
# pool_c3 <- colSums(d_c3)
# pool_c4 <- colSums(d_c4)
# pool_c5 <- colSums(d_c5)
# 
# # Combine them back into a single matrix where each row is a Cycle
# pooled_otu <- rbind(pool_c2, pool_c3, pool_c4, pool_c5)
# rownames(pooled_otu) <- c("C2", "C3", "C4", "C5")

# rarecurve(pooled_otu, 
#           step = 40, 
#           lwd = 2, 
#           label = TRUE,         
#           xlab = "Sample size", 
#           ylab = "Number of ASVs")

```

```{r}
# save plot
graphics.off()
pdf("rarefaction_combined.pdf", width = 8, height = 10) # Made it taller for 2 plots
par(mfrow = c(2, 1))
rarecurve(otu, 
          step = 40, 
          label = FALSE, 
          xlab = "Sample size", 
          ylab = "Number of ASVs",
          main = "All samples")

rarecurve(otu_sub, 
          step = 40, 
          label = FALSE, 
          xlab = "Sample size", 
          ylab = "Number of ASVs",
          main = "Live trap samples")

# rarecurve(pooled_otu, 
#           step = 40, 
#           lwd = 2, 
#           label = TRUE, 
#           xlab = "Sample size", 
#           ylab = "Number of ASVs",
#           main = "Pooled by cycles for live trap samples",
#           xlim = c(0, max(rowSums(pooled_otu)) * 1.1))

dev.off()

```

## NMDS
```{r}

dist <- avgdist(otu, sample = 1066, dmethod = "bray") #rarify to control for uneven number of reads per sample. lowest number of reads in a sample is 1066.

nmds_sub<- metaMDS(dist)

nmds_sample_sub <- scores(nmds_sub) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

goodness(nmds_sub)
stressplot(nmds_sub)

nmds_sub$stress

#plot by sample type
nmds_sample <- nmds_sample_sub %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA-Sc", "SA1","SA2" )) %>%
    mutate(sample = fct_relevel(sample, "water_CTD", "water_pump", "trap_fix", "trap_live", "core")) %>%
    ggplot() +
    stat_ellipse(aes(x=NMDS1, y= NMDS2, colour = sample)) +
    geom_point(aes(x=NMDS1, y= NMDS2, colour = sample, shape = cycle_name_2
                   ), size =2.5) +
    scale_colour_manual(values = sample_palette) +
    scale_shape_manual(values=c(3,7,16,17,15)) + # shapes for cycle
    theme(legend.position = "right",
        axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=13),
        axis.text.y = element_text(size = 10),
        axis.title.y =element_text(size=13)) +
    annotate("text", x = 0.3, y = -0.5, label = "Stress = 0.20", size = 5)


nmds_sample

# pdf("nmds_samples.pdf", height=5, width=6.5) ; plot(nmds_sample) ; dev.off()
```


```{r}
set.seed(1)
# dist.bray <- avgdist(otu, sample = 2446, dmethod = "bray") #rarify to control for uneven number of reads per sample.
# nmds<- metaMDS(dist.bray)
asv_hel <- decostand(otu, "hellinger") #raw counts, not transformed
nmds <- metaMDS(asv_hel, distance = "bray", autotransform = FALSE)
# nmds<- metaMDS(otu, distance ="bray")

goodness(nmds)
stressplot(nmds)

nmds$stress

nmds_sample <- as.data.frame(scores(nmds)$sites) %>%
  as.tibble(rownames="sample_name") %>%
  inner_join(., sample, by = "sample_name")

#plot by sample type
nmds_sample <- nmds_sample_sub %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA-Sc", "SA1","SA2" )) %>%
    mutate(sample = fct_relevel(sample, "water_CTD", "water_pump", "trap_fix", "trap_live", "core")) %>%
    ggplot() +
    stat_ellipse(aes(x=NMDS1, y= NMDS2, colour = sample)) +
    geom_point(aes(x=NMDS1, y= NMDS2, colour = sample, shape = cycle_name_2
                   ), size =2.5) +
    scale_colour_manual(values = sample_palette) +
    scale_shape_manual(values=c(3,7,16,17,15)) + # shapes for cycle
    theme(legend.position = "right",
        axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=13),
        axis.text.y = element_text(size = 10),
        axis.title.y =element_text(size=13)) +
    annotate("text", x = 0.3, y = -0.5, label = "Stress = 0.20", size = 5)


nmds_sample


```

## PERMANOVA

```{r}
dist <- vegdist(asv_hel, method = "bray")
dist_df <- as.data.frame(as.matrix(dist))
dist_df$sample_name <-rownames(dist_df) #row names to column
rownames(dist_df) <- c() #remove row names
meta_dist <-  inner_join(sample, dist_df, by = "sample_name") #join with the metdata
all_dist <- dist_df %>%
    select(all_of(.[["sample_name"]])) %>%
    as.dist() #something to change to a distance matrix for adonis

adonis2(all_dist~sample, data=meta_dist)
adonis2(all_dist~water_mass, data=meta_dist)

pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$water_mass)
pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$sample)
# adonis2(all_dist~sample+cycle_name_2, data=meta_dist)
# pairwiseAdonis::pairwise.adonis(all_dist, factors = meta_dist$cycle_name_2)


# betadisper analysis - sample type
bd <- betadisper(all_dist, meta_dist$sample)
anova(bd)
permutest(bd)
permutest(bd, pairwise = TRUE) # pairwise betadisper indicates sw monsoon is significantly more homogenous 



# betadisper analysis - water mass
bd <- betadisper(all_dist, meta_dist$water_mass)
anova(bd)
permutest(bd)


# meta_dist %>% count(sample)

# # Tukey's Honest Significant Differences
# (mod.HSD <- TukeyHSD(bd))
# plot(mod.HSD)

# meta_dist %>% group_by(sample) %>% count(sample)
```

```{r}
# 1. add pseudocount
otu_table(ps) <- otu_table(ps) + 1

# 2. CLR transform
physeq_clr <- microbiome::transform(ps, "clr")

# 3. extract matrix
clr_matrix <- t(otu_table(physeq_clr))

# 4. PCA
pca <- prcomp(clr_matrix, scale. = FALSE)

# variance explained
var_exp <- (pca$sdev^2 / sum(pca$sdev^2)) * 100

# 5. build plotting dataframe
meta <- data.frame(sample_data(ps))

pca_df <- data.frame(
  PC1 = pca$x[,1],
  PC2 = pca$x[,2],
  sample = meta$sample,
  water_mass = meta$water_mass
)

# 6. compute centroids
centroids <- pca_df %>%
  group_by(sample) %>%
  summarise(
    PC1 = mean(PC1),
    PC2 = mean(PC2)
  )

# 7. Aitchison distance
aitchison <- dist(clr_matrix)

# 8. PERMANOVA
perm <- adonis2(aitchison ~ sample, data = meta)
pval <- perm$`Pr(>F)`[1]

# 9. PCA plot
ggplot(pca_df, aes(PC1, PC2, color = sample)) +

  geom_point(size = 3, alpha = 0.9) +

  # confidence ellipses
  stat_ellipse(level = 0.95, linewidth = 1) +

  # centroids
  geom_point(
    data = centroids,
    aes(PC1, PC2, fill = sample),
    shape = 23,
    size = 5,
    color = "black"
  ) +

  theme_classic() +

  labs(
    title = "PCA of CLR-transformed community composition",
    subtitle = paste("PERMANOVA p =", signif(pval,3)),
    x = paste0("PC1 (", round(var_exp[1],1), "%)"),
    y = paste0("PC2 (", round(var_exp[2],1), "%)")
  )


# 6. Build dataframe for plotting
meta <- data.frame(sample_data(ps))

pca_df <- data.frame(
  PC1 = pca$x[,1],
  PC2 = pca$x[,2],
  PC3 = pca$x[,3],
  group = meta$sample
)

# 7. Compute centroids
centroids <- pca_df %>%
  group_by(group) %>%
  summarise(
    PC1 = mean(PC1),
    PC2 = mean(PC2),
    PC3 = mean(PC3)
  )

# 8. Aitchison distance for PERMANOVA
aitchison <- dist(clr_matrix)
perm <- adonis2(aitchison ~ sample, data = meta)
pval <- perm$`Pr(>F)`[1]

# Function to make PCA plot
plot_pca <- function(df, xvar, yvar, var_exp, centroids, pval){
  ggplot(df, aes_string(x = xvar, y = yvar, color = "group")) +
    geom_point(size = 3, alpha = 0.9) +
    stat_ellipse(level = 0.95, linewidth = 1) +
    geom_point(
      data = centroids,
      aes_string(x = xvar, y = yvar, fill = "group"),
      shape = 23,
      size = 5,
      color = "black"
    ) +
    theme_classic() +
    labs(
      title = paste0(xvar, " vs ", yvar, " PCA"),
      subtitle = paste("PERMANOVA p =", signif(pval,3)),
      x = paste0(xvar, " (", round(var_exp[as.numeric(substr(xvar,3,3))],1), "%)"),
      y = paste0(yvar, " (", round(var_exp[as.numeric(substr(yvar,3,3))],1), "%)")
    )
}

# 9. Generate three plots
p1 <- plot_pca(pca_df, "PC1", "PC2", var_exp, centroids, pval)
# p2 <- plot_pca(pca_df, "PC1", "PC3", var_exp, centroids, pval)
# p3 <- plot_pca(pca_df, "PC2", "PC3", var_exp, centroids, pval)

# 10. Display plots
p1
# p2
# p3
```


