---
title: "Shared community V3"
author: "Denise Ong"
date: "2024-04-15"
output: html_document
---
Comparisons between sample types

1. Surface vs lower water column, fixed traps and sediment - proportion of reads plot and ASV richness table. Separate between ST and SA only between proportion of reads. 

separating pump exported and not exported

2. Fixed vs live based on ASV presence absence - ASV richness table

# Initialize
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(here)

source(here("init_files", "init.R")) #for libraries
source(here("init_files", "init_markdown.R")) #for markdown libraries
source(here("init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps

set.seed(1)

# Subdivision palette
divisionPalette <- readxl::read_excel("init_files/division_palette.xlsx")
divisionPalette <- structure(divisionPalette$colour,.Names=divisionPalette$subdivision) 
```

# 1. Exported community sources and sinks

Final options:

- separate by water mass
- separate analysis for dinos.

## load data
```{r}
source(here("init_files", "read_data_v3.R")) #load phyloseq and modify variables
ps_all
source(here("init_files", "filt_samples.R")) #filter out replicates, blanks etc. 
ps
ps_filt <- ps %>% 
  subset_taxa(subdivision != "Dinoflagellata") # edit here based on the analysis conducted
ps_filt


ps_norm <- phyloseq_normalize_median(ps_filt)
ps_long <- phyloseq_transform_to_long(ps_norm)
```

## assign ASVs
```{r}
# separate by water mass
ps_long_common <- data.frame()
for (i in c("ST", "SA")) {
  
  # filter options here - only include very surface water col - depth cat 1-3 roughly 5-30 m. And separate analysis by water mass.
  ps_long_sub <- ps_long %>%
    filter(sample %in% c("trap_fix", "water_CTD", "core", "water_pump")) %>%
    filter(water_mass == i) 
  
  upset_df <- ps_long_sub %>%
    select(asv_code,sample) %>% # group by the level of classification
    distinct() %>%
    mutate(n=1) %>%
    pivot_wider(names_from = sample,
                values_fill = 0,
                values_from = n) %>%
    data.frame() # this part is important at the end of forming the df if not will have error later
  
  taxa <- as.data.frame(ps_filt@tax_table) %>%
    unique()
  
  # proportion of shared ASVs in each sample type
  common_asvs <- upset_df %>%
    mutate(shared_asvs = case_when(trap_fix ==1 & water_CTD ==1 & core ==1 & water_pump == 1 ~"surface exported",
                                   trap_fix ==1 & water_CTD ==1 & core ==1 & water_pump == 0 ~"surface exported",
                                   trap_fix ==1 & water_CTD ==1 & water_pump == 1 & core==0 ~ "surface not exported",
                                   trap_fix ==1 & water_CTD ==1 & water_pump == 0 & core==0 ~ "surface not exported",
                                    trap_fix ==1 & water_CTD ==0 & core ==1 & water_pump == 1 ~"pump exported",
                                    trap_fix ==1 & water_CTD ==0 & core ==0 & water_pump == 1 ~"pump not exported",
                                    trap_fix ==1 & water_CTD ==0 & core ==0 & water_pump == 0 ~"unique",
                                    trap_fix ==0 & water_CTD ==1 & core ==0 & water_pump == 0 ~"unique",
                                    trap_fix ==0 & water_CTD ==0 & core ==1 & water_pump == 0 ~"unique",
                                    trap_fix ==0 & water_CTD ==0 & core ==0 & water_pump == 1 ~"unique",
                                    trap_fix ==0 & water_CTD ==1 & core ==0 & water_pump == 1 ~"unique",
                                    trap_fix ==0 & water_CTD ==0 & core ==1 & water_pump == 1 ~"unique",
                                    TRUE~ "others"))
  ps_long_common_sub <- ps_long_sub %>%
    full_join(common_asvs) # %>%
    # filter(!(sample == "water_pump"))
  
   ps_long_common <- bind_rows(ps_long_common, ps_long_common_sub) %>%
     filter(shared_asvs != "others") # remove the other combinations, takes up 0.9 to 3.2% of reads in the samples
}


check <- ps_long_common %>%
  select(depth, cycle_name_2, sample) %>%
  unique()

```

## checks
### Visualise community as a whole
not using as a main plot, only for checking the community
```{r,eval=FALSE}
p <- ps_long_common %>%
  mutate(shared_asvs = case_when(
                                 # shared_asvs == "surface not exported" ~ "surface not exported",
                                 # shared_asvs == "pump not exported" ~ "pump not exported",
                                 shared_asvs == "surface exported" ~ "surface exported",
                                 # shared_asvs == "pump exported" ~ "pump exported",
                                 TRUE ~ "not shared")) %>%
  mutate(subdivision = case_when(shared_asvs == "not shared" ~ "not shared",
                                 shared_asvs == "surface exported" ~ subdivision,
                                 # shared_asvs == "pump exported" ~ subdivision
                                 # shared_asvs == "surface not exported" ~ subdivision,
                                 # shared_asvs == "pump not exported" ~ subdivision
                                 )) %>%
  # filter(shared_asvs == "surface and pump") %>%
  # filter(sample !="core") %>%
  mutate(sample = fct_relevel(sample, "water_CTD", "water_pump", "trap_fix", "core")) %>%
  # filter(sample %in% c("trap_fix", "core")) %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA2", "SA1", "SA-Sc", "ST2", "ST1")) %>%
  mutate(subdivision = case_when(subdivision == "Ciliophora" ~ "Ciliophora",
                                 subdivision == "Haptophyta_X" ~ "Haptophyta_X",
                                 subdivision == "Bigyra" ~ "Bigyra",
                                 subdivision == "Prasinodermophyta_X" ~ "Prasinodermophyta_X",
                                 subdivision == "Gyrista" ~ "Gyrista",
                                 subdivision == "Radiolaria" ~ "Radiolaria",
                                 subdivision == "Chlorophyta_X" ~ "Chlorophyta_X",
                                 class == "Syndiniales" ~ "Syndiniales",
                                 class == "Dinophyceae" ~ "Dinophyceae",
                                 subdivision == "not shared" ~ "not shared",
                                 TRUE ~"Others")) %>%
    mutate(subdivision = fct_relevel(subdivision, "not shared", "Others", "Bigyra", "Ciliophora", "Radiolaria", "Gyrista", "Haptophyta_X","Prasinodermophyta_X","Chlorophyta_X"
                                     , "Syndiniales", "Dinophyceae"
                                     )) %>%
  # mutate(subdivision = fct_rev(fct_infreq(subdivision))) %>%
  ggplot(aes(x=n_reads, y=cycle_name_2, color=subdivision, fill=subdivision)) + 
        geom_bar(position="fill", stat="identity") + 
        theme(panel.spacing.y = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              strip.text.y.right = element_text(angle = 0))+
        facet_grid(.~sample, scales = "free_y", space = "free")+
        scale_color_manual(values=divisionPalette) +
        scale_fill_manual(values=divisionPalette) +
        xlab("Percent of reads")+
        ylab("Cycle")+
  ggtitle("Surface and pump sources exported to the sediments")
p

# pdf("supp_shared_test.pdf", height= 3, width=7) ; plot(p2) ; dev.off()

```
For dinoflagellata
```{r}

p3 <- ps_long_common %>%
  mutate(shared_asvs = case_when(
                                 # shared_asvs == "surface not exported" ~ "surface not exported",
                                 # shared_asvs == "pump not exported" ~ "pump not exported",
                                 shared_asvs == "surface exported" ~ "surface exported",
                                 shared_asvs == "pump exported" ~ "pump exported",
                                 TRUE ~ "not shared")) %>%
  mutate(class = case_when(shared_asvs == "not shared" ~ "not shared",
                                 shared_asvs == "surface exported" ~ class,
                                 shared_asvs == "pump exported" ~ class
                                 # shared_asvs == "surface not exported" ~ class,
                                 # shared_asvs == "pump not exported" ~ class
                                 )) %>%
  mutate(sample = fct_relevel(sample, "water_CTD", "water_pump", "trap_fix")) %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA2", "SA1", "SA-Sc", "ST2", "ST1")) %>%
  mutate(class = case_when(class == "Syndiniales" ~ "Syndiniales",
                           class == "Dinophyceae" ~ "Dinophyceae",
                           class == "not shared" ~ "not shared",
                           TRUE ~"Others")) %>%
    mutate(class = fct_relevel(class, "not shared", "Others", "Syndiniales", "Dinophyceae")) %>%
  ggplot(aes(x=n_reads, y=cycle_name_2, color=class, fill=class)) + 
        geom_bar(position="fill", stat="identity") + 
        theme(panel.spacing.y = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              strip.text.y.right = element_text(angle = 0))+
        facet_grid(.~sample, scales = "free_y", space = "free")+
        scale_color_manual(values=divisionPalette) +
        scale_fill_manual(values=divisionPalette) +
        xlab("Percent of reads")+
        ylab("Cycle")+
  ggtitle("Surface and pump")
p3

```


### summary for results
```{r}
# proportion for each shared asv combination
otu_rel_abund <- ps_long_common %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>% # change reads to proportion
    ungroup() %>%
    select(-n_reads) %>%
    group_by(file_code, shared_asvs) %>%
    mutate(rel_abund = 100*sum(rel_abund)) %>% # add percent of reads for each sample combination tgt
    select(water_mass, cycle_name_2, sample, file_code, shared_asvs, rel_abund) %>%
    distinct() %>%
    pivot_wider(names_from = shared_asvs,
                values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= "surface exported":"pump exported",
                 names_to = "shared_asvs",
                 values_to = "rel_abund")

summary <- otu_rel_abund %>%
  group_by(sample, shared_asvs) %>%
  summarise(mean = round(mean(rel_abund), 1),
            sd = round(sd(rel_abund), 1)) #%>%
  # filter(mean !=0) %>%
  # unite("mean_sd", mean:sd, sep = "±") 
summary 

```

```{r}
#rel abund compare between upper and lower water column ASVs in traps
otu_rel_abund <- ps_long_common %>%
  # filter(shared_asvs %in% c("unique")) %>% #options to filer for each type of shared ASV
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:ASV,
                 names_to = "level",
                 values_to = "taxon") 

taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "class") %>% #choose level here
    group_by(water_mass, sample, cycle_name_2, file_code, shared_asvs, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !water_mass:shared_asvs,
                 names_to = "taxon",
                 values_to = "rel_abund")

class_df <- ps_long %>%
    select(division:class) %>%
    distinct()

# ## division level
summary <- taxon_rel_abund %>%
    # filter(cycle_name_2 == "SA2") %>%
  group_by(sample,shared_asvs,taxon) %>%
  summarise(#median = median(rel_abund),
            mean = round(mean(rel_abund), 1),
            sd = round(sd(rel_abund), 1)) %>%
  filter(mean !=0)  %>%
  # unite("mean_sd", mean:sd, sep = "±") %>%
  left_join(class_df,by=c("taxon" = "class"))
summary
```

```{r}
# proportion of each taxa
otu_rel_abund <- ps_long_common %>%
  # filter(shared_asvs %in% c("all", "water and trap")) %>% #options to filer for each type of shared ASV
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:ASV,
                 names_to = "level",
                 values_to = "taxon") 

taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "subdivision") %>% #choose level here
    group_by(water_mass, sample, cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

class_df <- ps_long %>%
    select(division:subdivision) %>%
    distinct()

# ## division level
summary <- taxon_rel_abund %>%
    # filter(cycle_name_2 == "SA2") %>%
  group_by(sample,water_mass,taxon) %>%
  summarise(#median = median(rel_abund),
            mean = round(mean(rel_abund), 1),
            sd = round(sd(rel_abund), 1)) %>%
  filter(mean !=0) #%>%
  # unite("mean_sd", mean:sd, sep = "±") %>%
  # left_join(class_df,by=c("taxon" = "class"))
summary 
```


## Plots
```{r}
library(viridis)
p1 <- ps_long_common %>%
  mutate(sample = fct_relevel(sample, "water_CTD", "trap_fix", "core")) %>%
  filter(sample %in% c("trap_fix", "core")) %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA2", "SA1", "SA-Sc", "ST2", "ST1")) %>%
  mutate(shared_asvs = fct_relevel(shared_asvs, "unique", "pump not exported", "pump exported", "surface not exported", "surface exported")) %>%
  ggplot(aes(x=n_reads, y=cycle_name_2, color=shared_asvs, fill=shared_asvs)) + 
        geom_bar(position="fill", stat="identity") + 
        theme(panel.spacing.y = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              strip.text.y.right = element_text(angle = 0))+
        facet_grid(.~sample, scales = "free_y", space = "free")+
        #   scale_colour_manual(values = c( "#FFFF99", "#FF9900", "#6BAED6", "#045A8D", "black"))+
        # scale_fill_manual(values = c( "#FFFF99", "#FF9900", "#6BAED6", "#045A8D", "black")) +
        scale_colour_manual(values = c("#777777", "#FFFF99", "#FF9900", "#6BAED6", "#045A8D", "black"))+
        scale_fill_manual(values = c("#777777", "#FFFF99", "#FF9900", "#6BAED6", "#045A8D", "black")) +
        xlab("Percent of reads")+
        ylab("Cycle")
p1

summary <- ps_long_common %>%
  select(domain:ASV, shared_asvs) %>%
  unique()

# pdf("shared_community_presentation_no dinos.pdf", height= 4, width=8) ; plot(p1) ; dev.off()

```

### Supplementary figure - shared ASVs by depth
checks for depth category
```{r}
library(viridis)
p6 <- ps_long_common %>%
  mutate(sample = fct_relevel(sample, "water_CTD", "trap_fix", "core")) %>%
  filter(sample %in% c("trap_fix")) %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA-Sc", "SA1", "SA2")) %>%
  mutate(shared_asvs = fct_relevel(shared_asvs, "unique", "pump not exported", "pump exported", "surface not exported", "surface exported")) %>%
  ggplot(aes(x=n_reads, y=depth_category, color=shared_asvs, fill=shared_asvs)) + 
        geom_bar(position="fill", stat="identity") + 
        theme(panel.spacing.y = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              strip.text.y.right = element_text(angle = 0))+
        facet_grid(.~cycle_name_2, scales = "free_y", space = "free")+
        scale_colour_manual(values = c("#777777", "#FFFF99", "#FF9900", "#6BAED6", "#045A8D", "black"))+
        scale_fill_manual(values = c("#777777", "#FFFF99", "#FF9900", "#6BAED6", "#045A8D", "black")) +
        xlab("Percent of reads")+
        ylab("Depth category") +
        ggtitle("Protist ASVs excluding Dinoflagellata")
p6

p <- p6/p7  + plot_annotation(tag_levels = 'A')  + plot_layout(guides = 'collect') & theme(plot.title=element_text(size=12))
p
pdf("supp_common asvs_depth.pdf", height= 5, width=9) ; plot(p) ; dev.off()

```
common asvs across all
```{r}
p2 <- ps_long_common %>%
  filter(shared_asvs == "surface exported") %>%
  # filter(sample !="water_CTD") %>%
  mutate(sample = fct_relevel(sample, "water_CTD", "trap_fix", "core")) %>%
  filter(sample %in% c("trap_fix", "core")) %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA2", "SA1", "SA-Sc", "ST2", "ST1")) %>%
  mutate(subdivision = case_when(subdivision == "Ciliophora" ~ "Ciliophora",
                                 subdivision == "Haptophyta_X" ~ "Haptophyta_X",
                                 subdivision == "Bigyra" ~ "Bigyra",
                                 subdivision == "Prasinodermophyta_X" ~ "Prasinodermophyta_X",
                                 subdivision == "Gyrista" ~ "Gyrista",
                                 subdivision == "Radiolaria" ~ "Radiolaria",
                                 subdivision == "Chlorophyta_X" ~ "Chlorophyta_X",
                                 class == "Syndiniales" ~ "Syndiniales",
                                 class == "Dinophyceae" ~ "Dinophyceae",
                                 TRUE ~"Others")) %>%
    mutate(subdivision = fct_relevel(subdivision, "Others", "Bigyra", "Ciliophora", "Radiolaria", "Gyrista", "Haptophyta_X", "Prasinodermophyta_X", "Chlorophyta_X", "Syndiniales", "Dinophyceae")) %>%
  ggplot(aes(x=n_reads, y=cycle_name_2, color=subdivision, fill=subdivision)) + 
        geom_bar(position="fill", stat="identity") + 
        theme(panel.spacing.y = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              strip.text.y.right = element_text(angle = 0)#,
              # legend.position = "none"
              )+
        facet_grid(.~sample, scales = "free_y", space = "free") +
        scale_color_manual(values=divisionPalette) +
        scale_fill_manual(values=divisionPalette) +
        xlab("Percent of reads")+
        ylab("Cycle")+
  ggtitle("Upp. exported")
p2
# pdf("presentation_exported.pdf", height= 4, width=6.5) ; plot(p2) ; dev.off()

```

what is not detected in the sediment
```{r}
p3 <- ps_long_common %>%
  filter(shared_asvs == "surface not exported") %>%
  # filter(sample != "water_CTD") %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA2", "SA1", "SA-Sc", "ST2", "ST1")) %>%
  filter(sample %in% c("trap_fix")) %>%
  mutate(sample = fct_relevel(sample, "water_CTD", "trap_fix", "core")) %>%
    mutate(subdivision = case_when(subdivision == "Ciliophora" ~ "Ciliophora",
                                 subdivision == "Haptophyta_X" ~ "Haptophyta_X",
                                 subdivision == "Bigyra" ~ "Bigyra",
                                 subdivision == "Prasinodermophyta_X" ~ "Prasinodermophyta_X",
                                 subdivision == "Gyrista" ~ "Gyrista",
                                 subdivision == "Radiolaria" ~ "Radiolaria",
                                 subdivision == "Chlorophyta_X" ~ "Chlorophyta_X",
                                 class == "Syndiniales" ~ "Syndiniales",
                                 class == "Dinophyceae" ~ "Dinophyceae",
                                 TRUE ~"Others")) %>%
    mutate(subdivision = fct_relevel(subdivision, "Others", "Bigyra", "Ciliophora", "Radiolaria", "Gyrista", "Haptophyta_X", "Prasinodermophyta_X", "Chlorophyta_X", "Syndiniales", "Dinophyceae")) %>%
  
  mutate(subdivision = case_when(subdivision == "Ciliophora" ~ "Ciliophora",
                                 subdivision == "Gyrista" ~ "Gyrista",
                                 subdivision == "Radiolaria" ~ "Radiolaria",
                                 class == "Syndiniales" ~ "Syndiniales",
                                 class == "Dinophyceae" ~ "Dinophyceae",
                                 TRUE~"Others"))%>%
  mutate(subdivision = fct_relevel(subdivision, "Others", "Gyrista", "Radiolaria", "Ciliophora", "Syndiniales", "Dinophyceae")) %>%
  ggplot(aes(x=n_reads, y=cycle_name_2, color=subdivision, fill=subdivision)) + 
        geom_bar(position="fill", stat="identity") + 
        theme(panel.spacing.y = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              strip.text.y.right = element_text(angle = 0),
              legend.position = "none"
              )+
        facet_grid(.~sample, scales = "free_y", space = "free") +
        scale_color_manual(values=divisionPalette) + 
        scale_fill_manual(values=divisionPalette) +
        xlab("Percent of reads")+
        ylab("Cycle") +
  ggtitle("Upp. not exported")
p3

# pdf("presentation_not exported.pdf", height= 4, width=4.5) ; plot(p3) ; dev.off()

```

what is exported from the lower water column
```{r}
p4 <- ps_long_common %>%
  filter(shared_asvs == "pump exported") %>%
  filter(sample != "water_pump") %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA2", "SA1", "SA-Sc", "ST2", "ST1")) %>%
  mutate(sample = fct_relevel(sample, "trap_fix", "core")) %>%
  mutate(subdivision = case_when(subdivision == "Ciliophora" ~ "Ciliophora",
                                 subdivision == "Gyrista" ~ "Gyrista",
                                 subdivision == "Radiolaria" ~ "Radiolaria",
                                 subdivision == "Bigyra" ~ "Bigyra",
                                 subdivision == "Cercozoa" ~ "Cercozoa",
                                 class == "Syndiniales" ~ "Syndiniales",
                                 class == "Dinophyceae" ~ "Dinophyceae",
                                 TRUE~"Others"))%>%
  mutate(subdivision = fct_relevel(subdivision, "Others", "Bigyra", "Ciliophora", "Radiolaria","Cercozoa", "Gyrista", "Syndiniales", "Dinophyceae")) %>%
  ggplot(aes(x=n_reads, y=cycle_name_2, color=subdivision, fill=subdivision)) + 
        geom_bar(position="fill", stat="identity") + 
        theme(panel.spacing.y = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              strip.text.y.right = element_text(angle = 0),
              # legend.position = "none"
              )+
        facet_grid(.~sample, scales = "free_y", space = "free")+
        scale_color_manual(values=divisionPalette) + 
        scale_fill_manual(values=divisionPalette) +
        xlab("Percent of reads")+
        ylab("Cycle")+
  ggtitle("Lower exported")
p4

# pdf("presentation_not exported.pdf", height= 4, width=4.5) ; plot(p3) ; dev.off()

```
what is not exported from the lower water column
```{r}
p5 <- ps_long_common %>%
  filter(shared_asvs == "pump not exported") %>%
  filter(sample != "water_pump") %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA2", "SA1", "SA-Sc", "ST2", "ST1")) %>%
  mutate(sample = fct_relevel(sample, "water_pump", "trap_fix", "core")) %>%
  mutate(subdivision = case_when(subdivision == "Ciliophora" ~ "Ciliophora",
                                 subdivision == "Gyrista" ~ "Gyrista",
                                 subdivision == "Radiolaria" ~ "Radiolaria",
                                 subdivision == "Bigyra" ~ "Bigyra",
                                 subdivision == "Cercozoa" ~ "Cercozoa",
                                 class == "Syndiniales" ~ "Syndiniales",
                                 class == "Dinophyceae" ~ "Dinophyceae",
                                 TRUE~"Others"))%>%
  mutate(subdivision = fct_relevel(subdivision, "Others", "Bigyra", "Ciliophora", "Radiolaria","Cercozoa", "Gyrista", "Syndiniales", "Dinophyceae"))  %>%
  ggplot(aes(x=n_reads, y=cycle_name_2, color=subdivision, fill=subdivision)) + 
        geom_bar(position="fill", stat="identity") + 
        theme(panel.spacing.y = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              strip.text.y.right = element_text(angle = 0),
              legend.position = "none"
              )+
        facet_grid(.~sample, scales = "free_y", space = "free")+
        scale_color_manual(values=divisionPalette) + 
        scale_fill_manual(values=divisionPalette) +
        xlab("Percent of reads")+
        ylab("Cycle")+
  ggtitle("Lower not exported")
p5

# pdf("presentation_not exported.pdf", height= 4, width=4.5) ; plot(p3) ; dev.off()

```

### Main fig 6
```{r}
#main
j1 <- (p2|p3) + plot_annotation(tag_levels = 'A') + plot_layout(widths = c(2, 1)) + plot_layout(guides = 'collect') & theme(plot.title=element_text(size=12)) 
p <- p1/j1 + plot_annotation(tag_levels = 'A')  & theme(plot.title=element_text(size=12)) 
p

pdf("shared_community_v4_main.pdf", height=6, width=10) ; plot(p) ; dev.off()

```

### Supplementary figs
```{r}
#supp
p <- (p4|p5) + plot_annotation(tag_levels = 'A') + plot_layout(widths = c(2, 1)) + plot_layout(guides = 'collect') & theme(plot.title=element_text(size=12)) 
p

pdf("shared_community_v4_supp.pdf", height=3, width=9) ; plot(p) ; dev.off()


#dino
j1 <- (p2|p3) + plot_annotation(tag_levels = 'A') + plot_layout(widths = c(2, 1)) + plot_layout(guides = 'collect') & theme(plot.title=element_text(size=12)) 
j2 <- (p4|p5)+ plot_annotation(tag_levels = 'A') + plot_layout(widths = c(2, 1)) + plot_layout(guides = 'collect') & theme(plot.title=element_text(size=12)) 
p <- p1/j1/j2 + plot_annotation(tag_levels = 'A')  & theme(plot.title=element_text(size=12)) 
p

pdf("shared_community_v4_dino.pdf", height=9, width=10) ; plot(p) ; dev.off()

```

## Supplementary table S2 - No. of ASVs summary
```{r}

#upset only accounts for presence/absence
upset_df <- ps_long %>%
  filter(sample %in% c("trap_fix", "water_CTD", "core", "water_pump")) %>%
  select(asv_code,sample) %>% # group by the level of classification
  distinct() %>%
  mutate(n=1) %>%
  pivot_wider(names_from = sample,
              values_fill = 0,
              values_from = n) %>%
  data.frame() # this part is important at the end of forming the df if not will have error later

taxa <- as.data.frame(ps_filt@tax_table) %>%
  unique

# proportion of shared ASVs in each sample type
common_asvs <- upset_df %>%
    mutate(shared_asvs = case_when(trap_fix ==1 & water_CTD ==1 & core ==1 & water_pump == 1 ~"surface exported",
                                   trap_fix ==1 & water_CTD ==1 & core ==1 & water_pump == 0 ~"surface exported",
                                   trap_fix ==1 & water_CTD ==1 & water_pump == 1 & core==0 ~ "surface not exported",
                                   trap_fix ==1 & water_CTD ==1 & water_pump == 0 & core==0 ~ "surface not exported",
                                    trap_fix ==1 & water_CTD ==0 & core ==1 & water_pump == 1 ~"pump exported",
                                    trap_fix ==1 & water_CTD ==0 & core ==0 & water_pump == 1 ~"pump not exported",
                                    trap_fix ==1 & water_CTD ==0 & core ==0 & water_pump == 0 ~"unique",
                                    trap_fix ==0 & water_CTD ==1 & core ==0 & water_pump == 0 ~"unique",
                                    trap_fix ==0 & water_CTD ==0 & core ==1 & water_pump == 0 ~"unique",
                                    trap_fix ==0 & water_CTD ==0 & core ==0 & water_pump == 1 ~"unique",
                                    trap_fix ==0 & water_CTD ==1 & core ==0 & water_pump == 1 ~"unique",
                                    trap_fix ==0 & water_CTD ==0 & core ==1 & water_pump == 1 ~"unique",
                                    TRUE~ "others"))
common_asvs_join <- ps_long %>%
  filter(sample %in% c("trap_fix", "water_CTD", "core", "water_pump")) %>%
  full_join(common_asvs) %>%
  mutate(sample = fct_relevel(sample, "water_CTD","water_pump",  "trap_fix", "core")) %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA2", "SA1", "SA-Sc", "ST2", "ST1")) %>%
  mutate(shared_asvs = fct_relevel(shared_asvs,  "unique", "pump not exported","pump exported", "surface not exported", "surface exported"))

# table
summary <- common_asvs_join %>%
  select(sample, shared_asvs, subdivision, class, asv_code) %>%
  group_by(sample, shared_asvs, subdivision, class, asv_code) %>%
  unique() %>%
  group_by(sample, shared_asvs, subdivision, class) %>%
  summarise(asv_num = n()) %>%
  filter(shared_asvs != "unique") %>%
  filter(shared_asvs != "others") %>%
  ungroup() %>%
  filter(!((sample == "water_pump" & shared_asvs == "surface not exported")|
         (sample == "water_pump" & shared_asvs == "surface exported"))) %>%
  select(-sample) %>%
  unique() %>%
  pivot_wider(names_from = shared_asvs, values_from = asv_num, values_fill = 0)

summary2 <- common_asvs_join %>%
  select(sample, shared_asvs, subdivision, class, asv_code) %>%
  group_by(sample, shared_asvs, subdivision, class, asv_code) %>%
  unique() %>%
  group_by(sample, shared_asvs, subdivision, class) %>%
  summarise(asv_num = n()) %>%
  filter(shared_asvs == "unique") %>%
  ungroup() %>%
  select(-shared_asvs) %>%
  pivot_wider(names_from = sample, values_from = asv_num, values_fill = 0)

summary_asv <- full_join(summary, summary2) %>%
  select(subdivision: "pump exported", trap_fix) %>% # remove this for the rest of the samples unique ASV
  mutate_all(~replace(., is.na(.), 0)) %>%
  mutate(total = rowSums(across(where(is.numeric)))) %>%
  filter(total !=0) %>%
  arrange(desc(total)) %>%
  arrange(subdivision)
  # summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE)))
 writexl::write_xlsx(summary_asv, path = "shared_asv_num_v2.xlsx")
 
 #check exported taxa in sediments
 summary_asv_sediment <- full_join(summary, summary2) %>%
  select(subdivision: "pump exported", core) %>% # remove this for the rest of the samples unique ASV
  mutate_all(~replace(., is.na(.), 0)) %>%
  select(-'surface not exported') %>%
  select(-'pump not exported') %>%
  mutate(total = rowSums(across(where(is.numeric)))) %>%
  filter(total !=0) %>%
  arrange(desc(total)) %>%
  arrange(subdivision) 
```

# 2. fixed vs live ASV richness
```{r}
source(here("init_files", "read_data_v3.R")) #load phyloseq and modify variables
ps_all
source(here("init_files", "filt_samples.R")) #filter out replicates, blanks, low abund ASVs etc. 
ps

ps_norm <- phyloseq_normalize_median(ps)
ps_long <- phyloseq_transform_to_long(ps_norm)

library(UpSetR)

# number of distinct ASVs per sample type
sample_asv <- ps_long %>%
  filter(sample %in% c("trap_fix", "trap_live")) %>%
  select(asv_code, sample) %>%
  distinct() %>%
  group_by(sample) %>%
  summarise(num_asv = n())
sample_asv

taxa <- as.data.frame(ps_filt@tax_table) %>%
  unique

#upset only accounts for presence/absence
upset_df <- ps_long %>%
  filter(sample %in% c("trap_fix", "trap_live")) %>%
  select(asv_code,sample) %>% # group by the level of classification
  distinct() %>%
  mutate(n=1) %>%
  pivot_wider(names_from = sample,
              values_fill = 0,
              values_from = n) %>%
  data.frame() # this part is important at the end of forming the df if not will have error later

fig_upset_asv <- UpSetR::upset(upset_df, 
                                # nsets = 7,
                                empty.intersections = "on", 
                                # nintersects = 15,
                                mainbar.y.label = "Number of ASVs", 
                                sets.x.label = "ASVs", 
                                order.by = "freq", 
                                mb.ratio = c(0.6, 0.4))
fig_upset_asv
# need to manually save


upset_taxa <- upset_df %>%
  mutate(group = case_when(trap_fix == 1 & trap_live == 1 ~ "both",
                           trap_fix == 1 & trap_live  == 0 ~ "trap_fix",
                           trap_fix == 0 & trap_live  == 1 ~ "trap_live")) %>%
  left_join(taxa, by=join_by(asv_code == ASV)) %>%
  group_by(group, order) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  complete(group, order) %>%
  replace_na(list(n= 0))

subdivision_list <- taxa %>%select(subdivision, order) %>% unique()

#proportion of fixed trap ASVs found in live traps
df <- upset_taxa %>%
  replace_na(list(n= 0)) %>%
  pivot_wider(names_from = group, values_from = n) %>%
  # select(-trap_live) %>%
  mutate(fixed_total = trap_fix+both) %>%
  mutate(live_total = trap_live+both) %>%
  # mutate(live_overlap = fixed_total-trap_fix) %>%
  mutate(live_prop = round((both/fixed_total)*100, 1)) %>%
  mutate(unique_live = trap_live) %>%
  select(order, fixed_total, live_prop, unique_live) %>%
  left_join(subdivision_list, by=join_by(order)) %>%
  select(subdivision, order, live_prop, fixed_total, unique_live) %>%
  arrange(subdivision, desc(live_prop)) %>%
  filter(live_prop != 0) %>%
  filter(fixed_total!= 0)

df  
writexl::write_xlsx(df, "fix_vs_live_ASV.xlsx")
df %>%
  summarise(mean= round(mean(live_prop),2),
            sd = round(sd(live_prop),2))
  
```
