---
title: "ALDEx"
format: html
---

# init
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Load package
library(here)
library(purrr)
library(ALDEx2)

source(here("init_files", "init.R")) #for libraries
source(here("init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps

set.seed(1)


# Subdivision palette
divisionPalette <- readxl::read_excel("init_files/division_palette.xlsx", sheet = "Sheet1")
divisionPalette <- structure(divisionPalette$colour,.Names=divisionPalette$subdivision) 

dinoPalette <- readxl::read_excel("init_files/division_palette.xlsx", sheet = "Sheet2")
dinoPalette <- structure(dinoPalette$colour,.Names=dinoPalette$order) 

```

# ALDEx2 from phyloseq function
```{r}
phyloseq_run_aldex <- function(ps) {
  counts <- as(otu_table(ps), "matrix")
  
  # Ensure taxa are rows
  if (!taxa_are_rows(ps)) {
    counts <- t(counts)
  }
  
  meta <- sample_data(ps)
  conds <- as.character(meta$sample_depth)
  print(conds)
  
  x <- aldex.clr(counts, conds, mc.samples=128, verbose=F, gamma=0.20)
  x.tt <- aldex.ttest(x, hist.plot=F, paired.test=FALSE, verbose=FALSE)
  x.effect <- aldex.effect(x, CI=T, verbose=F, include.sample.summary=F, paired.test=FALSE)
  
  x.all <- data.frame(x.tt,x.effect) %>%
    rownames_to_column(var = "ASV")
  
  taxa <- as.data.frame(tax_table(ps))
  
  x.taxa <- full_join(x.all, taxa)
  
  return(x.taxa)
}
```

# load data
```{r}
source(here("init_files", "read_data_v3.R")) #load phyloseq and modify variables
ps_all
source(here("init_files", "filt_samples.R")) #filter out replicates, blanks etc. 
ps
```



# subset phyloseq
```{r}
deseq.subset <- ps %>%
  # combine depth 1+2, 3+4
  ps_filter(sample == "water_CTD" | sample == "trap_fix") %>%
  ps_filter(cycle_name != "SA-Sc-B") %>% # no fixed traps in this cycle
  ps_mutate(sample_depth = case_when(sample == "trap_fix" ~ paste(sample, depth_category, sep = "_"), 
                                     # add label for depth only for fixed traps
                                     sample == "water_CTD" ~ sample)) %>%
  ps_mutate(sample_depth = case_when(sample_depth == "trap_fix_1" ~ "trap_fix_1_2",
                                     sample_depth == "trap_fix_2" ~ "trap_fix_1_2",
                                     sample_depth == "trap_fix_3" ~ "trap_fix_3_4",
                                     sample_depth == "trap_fix_4" ~ "trap_fix_3_4",
                                     TRUE ~ as.character(sample_depth))) %>%
  ps_mutate(sample_depth_wm = paste(sample_depth, water_mass, sep = "_"))  %>%
  tax_filter(min_total_abundance = 11)

deseq.subset

sample <- as.data.frame(deseq.subset@sam_data)

```
# list of ASVs from sediment
```{r}
ps_core <- ps %>% 
   ps_filter(sample == "core")
ps_long <- phyloseq_transform_to_long(ps_core)

core_asvs <- ps_long %>%
  select(ASV, water_mass) %>%
  unique() %>%
  mutate(sediment = "detected")

core_asvs_ST <- core_asvs %>%
  filter(water_mass == "ST") %>%
  select(ASV, sediment)

core_asvs_SA <- core_asvs %>%
  filter(water_mass == "SA") %>%
  select(ASV, sediment)
```

# ST analysis
Subset ST water mass, filter ASVs for minimum prevalence 5%, and only asvs present in all sample groups.
```{r}
# filter for ASVs with minimum prevalence 10%
deseq.st <- deseq.subset %>%
  ps_filter(water_mass == "ST") %>%
  tax_filter(min_prevalence = 0.10)

# filter for asvs present in all the sample groups
group_var <- "sample_depth"

taxa_by_group <- sample_data(deseq.st)[[group_var]] %>% 
  unique() %>% 
  purrr::map(function(x) {
    # Subset to one group at a time
    ps_sub <- prune_samples(sample_data(deseq.st)[[group_var]] == x, deseq.st)
    # Get names of taxa that have a sum > 0 in this specific group
    phyloseq::taxa_names(phyloseq::prune_taxa(taxa_sums(ps_sub) > 0, ps_sub))
  })

common_taxa <- purrr::reduce(taxa_by_group, intersect)

deseq.st.filt <- prune_taxa(common_taxa, deseq.st)

deseq.st.filt

# no. of samples in each group
deseq.st.filt@sam_data %>%
  group_by(sample_depth_wm) %>%
  summarise(n=n())

```


run aldex analysis separately for export and transfer.
```{r}
ps <- deseq.st.filt %>%
  ps_filter(sample_depth != "trap_fix_3_4")

res_st.export<- phyloseq_run_aldex(ps) 
  
ps <- deseq.st.filt %>%
  ps_filter(sample_depth != "water_CTD")

res_st.transfer <- phyloseq_run_aldex(ps) 
```

plot 
```{r}
res_filt_st.export <- res_st.export %>%
  rename_with(~ paste0(.x, "_export"), .cols = we.ep:overlap) %>%
  mutate(is_significant_export = (overlap_export <= 0.4 & abs(effect_export) > 0.25)) %>%
  mutate(diff.btw_export = diff.btw_export * (-1)) 

res_filt_st.transfer <- res_st.transfer %>%
  rename_with(~ paste0(.x, "_transfer"), .cols = we.ep:overlap) %>%
  mutate(is_significant_transfer = (overlap_transfer <= 0.4 & abs(effect_transfer) > 0.25)) %>%
  select(ASV:overlap_transfer, is_significant_transfer)

res_filt_st.join <- full_join(res_filt_st.export, res_filt_st.transfer, by = join_by(ASV)) %>%
  full_join(core_asvs_ST) %>%
  mutate(sediment = coalesce(sediment, "not detected")) %>%
  filter(!(is.na(diff.btw_export))) 

writexl::write_xlsx(res_filt_st.join, "aldex_results_st.xlsx")
```

## plot

### supp - protist except dino
```{r}
all_classes <- res_filt_st.join %>%
  filter(subdivision != "Dinoflagellata") %>%
  filter(is_significant_export | is_significant_transfer) %>%
  distinct(class) %>%
  pull(class)

subdivision_list <- res_filt_st.join %>%
  filter(subdivision != "Dinoflagellata") %>%
  select(subdivision, class) %>%
  distinct()

g1 <- res_filt_st.join %>%
  filter(subdivision != "Dinoflagellata") %>%
  filter(is_significant_export) %>%
  complete(class = all_classes) %>%
  left_join(subdivision_list,by = "class") %>%
  ggplot(aes(x=fct_rev(class), y=diff.btw_export)) + 
  geom_point(aes(shape=sediment, size=sediment, fill=subdivision.y), colour="black", na.rm = TRUE) +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("Median CLR difference - Water CTD <--> Upper fixed traps") +
  xlab("Class")+
  ggtitle("Export potential") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(subdivision.y~., scales = "free", space = "free", switch = "y") +
  # scale_shape_manual(values=c(3, 16, 17)) +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g1
# pdf("deseq_legend.pdf", height= 10, width=7) ; plot(g1) ; dev.off()


g2 <- res_filt_st.join %>%
  filter(subdivision != "Dinoflagellata") %>%
  filter(is_significant_transfer) %>%
  complete(class = all_classes) %>%
  left_join(subdivision_list,by = "class") %>%
  # mutate(class = factor(class, levels = all_classes)) %>%
  ggplot(aes(x=fct_rev(class), y=diff.btw_transfer)) + 
  geom_point(aes(shape=sediment, size=sediment, fill=subdivision.y), colour="black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("Median CLR difference - Upper fixed traps <--> Lower fixed traps") +
  xlab("Class")+
  ggtitle("Transfer potential") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(subdivision.y~., scales = "free", space = "free", switch = "y") +
  # scale_shape_manual(values=c(3, 16, 17)) +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g2
# pdf("deseq_legend.pdf", height= 10, width=7) ; plot(g1) ; dev.off()

```

### supp - dino
```{r}
all_order <- res_filt_st.join %>%
  filter(subdivision == "Dinoflagellata") %>%
  filter(is_significant_export | is_significant_transfer) %>%
  distinct(order) %>%
  pull(order)

class_list <- res_filt_st.join %>%
  filter(subdivision == "Dinoflagellata") %>%
  select(class, order) %>%
  distinct()

g1_dino <- res_filt_st.join %>%
  filter(subdivision == "Dinoflagellata") %>%
  filter(is_significant_export) %>%
  complete(order = all_order) %>%
  left_join(class_list, by = "order") %>%
  ggplot(aes(x=fct_rev(order), y=diff.btw_export)) + 
  geom_point(aes(shape=sediment, size=sediment, fill=class.y), colour="black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("Median CLR difference - Water CTD <--> Upper fixed traps") +
  xlab("Class")+
  ggtitle("Export potential") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(class.y~., scales = "free", space = "free", switch = "y") +
  # scale_shape_manual(values=c(3, 16, 17)) +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g1_dino
# pdf("deseq_legend.pdf", height= 10, width=7) ; plot(g1) ; dev.off()


g2_dino <- res_filt_st.join %>%
  filter(subdivision == "Dinoflagellata") %>%
  filter(is_significant_transfer) %>%
  complete(order = all_order) %>%
  left_join(class_list, by = "order") %>%
  ggplot(aes(x=fct_rev(order), y=diff.btw_transfer)) + 
  geom_point(aes(shape=sediment, size=sediment, fill=class.y), colour="black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("Median CLR difference - Upper fixed traps <--> Lower fixed traps") +
  xlab("Order")+
  ggtitle("Transfer potential") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(class.y~., scales = "free", space = "free", switch = "y") +
  # scale_shape_manual(values=c(3, 16, 17)) +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g2_dino
# pdf("deseq_legend.pdf", height= 10, width=7) ; plot(g1) ; dev.off()

```

### main
```{r}

g3 <- res_filt_st.join %>% 
  filter(subdivision != "Dinoflagellata") %>%
  filter(is_significant_export & is_significant_transfer) %>%
  mutate(subdivision = fct_relevel(subdivision,
                                   "Chlorophyta_X", "Cryptophyta_X", "Gyrista", "Haptophyta_X", "Prasinodermophyta_X", "Ciliophora", "Radiolaria")) %>%
  ggplot(aes(x = diff.btw_export, y = diff.btw_transfer)) +
  geom_point(aes(shape=sediment, fill=subdivision), colour="black", size = 4) +
  geom_text_repel(
    data = res_filt_st.join %>% 
      filter(subdivision != "Dinoflagellata") %>%
      filter(is_significant_export & is_significant_transfer) %>% 
      filter(sediment == "detected"),
    aes(label = class, colour = subdivision),
    size = 3, max.overlaps = 30,
    box.padding = 0.25,
    point.padding = 0.3,
    hjust = "outward",
    direction = "y")+
  geom_vline(xintercept = 0, linetype="solid") +  
  geom_hline(yintercept = 0, linetype="solid") +  
  xlab("Median CLR difference - Water CTD <--> Upper fixed traps") +
  ylab("Median CLR difference - Upper <--> Lower fixed traps") +
  scale_fill_manual(values=divisionPalette) +
  scale_colour_manual(values=divisionPalette) +
  theme(axis.text.x = element_text(size = 9),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 9),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        # legend.position = "none"
        ) +
  scale_shape_manual(values = c(24,21))+
  coord_fixed() +
  ggtitle("Subtropical") +
  scale_y_continuous(limits = c(-10, 10))+
  scale_x_continuous(limits = c(-10, 10))

g3

```



```{r}

g4 <- res_filt_st.join %>% 
  filter(subdivision == "Dinoflagellata") %>%
  # filter(is_significant_transfer) %>%
  filter(is_significant_export & is_significant_transfer) %>%
  ggplot(aes(x = diff.btw_export, y = diff.btw_transfer)) +
  geom_point(aes(shape=sediment, size=sediment, fill=class, colour = class)) +
  geom_text_repel(
    data = res_filt_st.join %>% 
      filter(subdivision == "Dinoflagellata") %>%
      filter(is_significant_export & is_significant_transfer) %>% 
      filter(sediment == "detected"),
    aes(label = order, colour = order),
    size = 3, max.overlaps = 30,
    box.padding = 0.25,
    point.padding = 0.3,
    hjust = "outward",
    direction = "y")+
  geom_vline(xintercept = 0, linetype="solid") +  
  geom_hline(yintercept = 0, linetype="solid") +  
  xlab("Median CLR difference - Water CTD <--> Upper fixed traps") +
  ylab("Median CLR difference - Upper <--> Lower fixed traps") +
  scale_fill_manual(values=divisionPalette) +
  scale_colour_manual(values=divisionPalette)+
  theme(axis.text.x = element_text(size = 9),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 9),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        # legend.position = "none"
        ) +
  scale_size_manual(values=c(4, 4)) +
  scale_shape_manual(values = c(24,21))+
  coord_fixed() +
  ggtitle("Subtropical") +
  scale_y_continuous(limits = c(-10, 10))+
  scale_x_continuous(limits = c(-10, 10))

g4


```


## ST- ASV raw data supp fig
```{r}

target_asvs <- res_filt_st.join %>%
  filter(subdivision != "Dinoflagellata") %>%
  # filter(is_significant_transfer) %>%
  filter(is_significant_export & is_significant_transfer) %>%
  pull(ASV) 

ps_rel <- transform_sample_counts(deseq.st.filt, function(x) x / sum(x)) %>%
  psmelt() 

plot_df <- ps_rel %>%
  filter(OTU %in% target_asvs)  %>%
  mutate(sample_depth = factor(sample_depth, 
                               levels = c("water_CTD", "trap_fix_1_2", "trap_fix_3_4"))) %>%
  mutate(label = paste0(subdivision, " | ", genus, "\n", OTU))

 p <- ggplot(plot_df, aes(x = sample_depth, y = Abundance, fill = sample_depth)) +
  geom_boxplot(alpha = 0.2, outlier.shape = NA) + 
  geom_jitter(width = 0.2, shape = 21, size = 3, alpha = 0.75) +
  facet_wrap(~label, scales = "free_y") + 
  theme_bw() +
  ggtitle("Subtropical - Differentially abundant ASVs")+
  labs(y = "Relative Abundance", x = "Sample Depth")+
   theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # rotate x-axis labels
    legend.text = element_text(size = 12),
  legend.title = element_text(size = 14),
  legend.key.size = unit(1.5, "lines"))  # size of legend boxes
p
 # pdf("aldex_st_raw rel abund.pdf", height= 9, width=17) ; plot(p) ; dev.off()
```

# SA analysis
```{r}

deseq.sa <- deseq.subset %>%
  ps_filter(water_mass == "SA") %>%
  tax_filter(min_prevalence = 0.1)


group_var <- "sample_depth"

# create a list of asvs present in each group
taxa_by_group <- sample_data(deseq.sa)[[group_var]] %>% 
  unique() %>% 
  purrr::map(function(x) {
    ps_sub <- prune_samples(sample_data(deseq.sa)[[group_var]] == x, deseq.sa)
    taxa_names(prune_taxa(taxa_sums(ps_sub) > 0, ps_sub))
  })

common_taxa <- purrr::reduce(taxa_by_group, intersect) # find ASVs common in all lists

# prune taxa, only keep those asvs that are common
deseq.sa.filt <- prune_taxa(common_taxa, deseq.sa)

deseq.sa.filt

# #no. of samples in each group
deseq.sa.filt@sam_data %>%
  group_by(sample_depth_wm) %>%
  summarise(n=n())

```



```{r}
ps <- deseq.sa.filt %>%
  ps_filter(sample_depth != "trap_fix_3_4")

res_sa.export<- phyloseq_run_aldex(ps)
  
ps <- deseq.sa.filt %>%
  ps_filter(sample_depth != "water_CTD")

res_sa.transfer <- phyloseq_run_aldex(ps)

```

```{r}

res_filt_sa.export <- res_sa.export %>%
  mutate(is_significant_export = (overlap <= 0.4 & abs(effect) > 0.25)) %>%
  mutate(diff.btw_export = diff.btw * (-1))  %>%
  select(is_significant_export, diff.btw_export, ASV, domain:species)

res_filt_sa.transfer <- res_sa.transfer %>%
  mutate(is_significant_transfer = (overlap <= 0.4 & abs(effect) > 0.25)) %>%
  mutate(diff.btw_transfer = diff.btw)  %>%
  select(is_significant_transfer, diff.btw_transfer, ASV, domain:species)

res_filt_sa.join <- full_join(res_filt_sa.export, res_filt_sa.transfer %>% select(ASV, diff.btw_transfer, is_significant_transfer), by = join_by(ASV)) %>%
  full_join(core_asvs_SA) %>%
  mutate(sediment = coalesce(sediment, "not detected")) %>%
  filter(!(is.na(diff.btw_export))) 

writexl::write_xlsx(res_filt_sa.join, "aldex_results_sa.xlsx")

```

## plot

### supp - protist except dino
```{r}
all_classes <- res_filt_sa.join %>%
  filter(subdivision != "Dinoflagellata") %>%
  filter(is_significant_export | is_significant_transfer) %>%
  distinct(class) %>%
  pull(class)

subdivision_list <- res_filt_sa.join %>%
  filter(subdivision != "Dinoflagellata") %>%
  select(subdivision, class) %>%
  distinct()

g5 <- res_filt_sa.join %>%
  filter(is_significant_export) %>%
  filter(subdivision != "Dinoflagellata") %>%
  complete(class = all_classes) %>%
  left_join(subdivision_list,by = "class") %>%
  ggplot(aes(x=fct_rev(class), y=diff.btw_export)) + 
  geom_point(aes(shape=sediment, size=sediment, fill=subdivision.y), colour="black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("Median CLR difference - Water CTD <--> Upper fixed traps") +
  xlab("Class") +
  ggtitle("Export potential") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(subdivision.y ~., scales = "free", space = "free", switch = "y") +
  # scale_shape_manual(values=c(3, 16, 17)) +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g5
# pdf("deseq_legend.pdf", height= 10, width=7) ; plot(g1) ; dev.off()

g6 <- res_filt_sa.join %>%
  filter(is_significant_transfer) %>%
  filter(subdivision != "Dinoflagellata") %>%
  complete(class = all_classes) %>%
  left_join(subdivision_list, by = "class") %>%
  ggplot(aes(x=fct_rev(class), y=diff.btw_transfer)) + 
  geom_point(aes(shape=sediment, size=sediment, fill=subdivision.y), colour="black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("Median CLR difference - Upper fixed traps <--> Lower fixed traps") +
  xlab("Class") +
  ggtitle("Transfer potential") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(subdivision.y ~., scales = "free", space = "free", switch = "y") +
  # scale_shape_manual(values=c(3, 16, 17)) +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g6
# pdf("deseq_legend.pdf", height= 10, width=7) ; plot(g1) ; dev.off()



```

###supp - dino
```{r}
all_order <- res_filt_sa.join %>%
  filter(subdivision == "Dinoflagellata") %>%
  filter(is_significant_export | is_significant_transfer) %>%
  distinct(order) %>%
  pull(order)

class_list <- res_filt_sa.join %>%
  filter(subdivision == "Dinoflagellata") %>%
  select(class, order) %>%
  distinct()

g5_dino <- res_filt_sa.join %>%
  filter(is_significant_export) %>%
  filter(subdivision == "Dinoflagellata") %>%
  complete(order = all_order) %>%
  left_join(class_list, by = "order") %>%
  ggplot(aes(x=fct_rev(order), y=diff.btw_export)) + 
  geom_point(aes(shape=sediment, size=sediment, fill=class.y), colour="black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("Median CLR difference - Water CTD <--> Upper fixed traps") +
  xlab("Order") +
  ggtitle("Export potential") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(class.y ~., scales = "free", space = "free", switch = "y") +
  # scale_shape_manual(values=c(3, 16, 17)) +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g5_dino
# pdf("deseq_legend.pdf", height= 10, width=7) ; plot(g1) ; dev.off()


g6_dino <- res_filt_sa.join %>%
  filter(is_significant_transfer) %>%
  filter(subdivision == "Dinoflagellata") %>%
  complete(order = all_order) %>%
  left_join(class_list, by = "order") %>%
  ggplot(aes(x=fct_rev(order), y=diff.btw_transfer)) + 
  geom_point(aes(shape=sediment, size=sediment, fill=class.y), colour="black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("Median CLR difference - Upper fixed traps <--> Lower fixed traps") +
  xlab("Order") +
  ggtitle("Transfer potential") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(class.y ~., scales = "free", space = "free", switch = "y") +
  # scale_shape_manual(values=c(3, 16, 17)) +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g6_dino
# pdf("deseq_legend.pdf", height= 10, width=7) ; plot(g1) ; dev.off()


```

### main
```{r}

g7 <- res_filt_sa.join %>% 
  filter(subdivision != "Dinoflagellata") %>%
  filter(is_significant_export & is_significant_transfer) %>%
  mutate(subdivision = fct_relevel(subdivision,
                                   "Chlorophyta_X", "Gyrista", "Haptophyta_X", "Prasinodermophyta_X", "Cercozoa", "Choanoflagellata", "Ciliophora", "Radiolaria")) %>%
  ggplot(aes(x = diff.btw_export, y = diff.btw_transfer)) +
  geom_point(aes(shape=sediment, fill=subdivision), colour="black", size = 4) +
  geom_text_repel(
    data = res_filt_sa.join %>% 
      filter(!(subdivision %in% c("Dinoflagellata", "Cercozoa", "Choanoflagellata"))) %>%
      filter(is_significant_export & is_significant_transfer) %>% 
      filter(sediment == "detected"),
    aes(label = class, colour = subdivision),
    size = 3, max.overlaps = 30,
    box.padding = 0.25,
    point.padding = 0.3,
    hjust = "outward",
    direction = "y")+
  geom_vline(xintercept = 0, linetype="solid") +  
  geom_hline(yintercept = 0, linetype="solid") +  
  xlab("Median CLR difference - Water CTD <--> Upper fixed traps") +
  ylab("Median CLR difference - Upper <--> Lower fixed traps") +
  scale_fill_manual(values=divisionPalette) +
  scale_colour_manual(values=divisionPalette) +
  theme(axis.text.x = element_text(size = 9),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 9),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        # legend.position = "none"
        ) +
  scale_shape_manual(values = c(24,21))+
  coord_fixed() +
  ggtitle("Subantarctic") +
  scale_y_continuous(limits = c(-10, 10))+
  scale_x_continuous(limits = c(-10, 10))

g7
```

```{r}

g8 <- res_filt_sa.join %>% 
  filter(subdivision == "Dinoflagellata") %>%
  filter(is_significant_export & is_significant_transfer) %>%
  ggplot(aes(x = diff.btw_export, y = diff.btw_transfer)) +
  geom_point(aes(shape=sediment, size=sediment, fill=class, colour = class)) +
  geom_text_repel(
    data = res_filt_sa.join %>% 
      filter(subdivision == "Dinoflagellata") %>%
      filter(is_significant_export & is_significant_transfer) %>% 
      filter(sediment == "detected"),
    aes(label = order, colour = class),
    size = 3, max.overlaps = 30,
    box.padding = 0.25,
    point.padding = 0.3,
    hjust = "outward",
    direction = "y")+
  geom_vline(xintercept = 0, linetype="solid") +  
  geom_hline(yintercept = 0, linetype="solid") +  
  xlab("Median CLR difference - Water CTD <--> Upper fixed traps") +
  ylab("Median CLR difference - Upper <--> Lower fixed traps") +
  scale_fill_manual(values=divisionPalette) +
  scale_colour_manual(values=divisionPalette)+
  theme(axis.text.x = element_text(size = 9),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 9),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        # legend.position = "none"
        ) +
  scale_size_manual(values=c(4, 4)) +
  scale_shape_manual(values = c(24,21))+
  coord_fixed() +
  ggtitle("Subantarctic") +
  scale_y_continuous(limits = c(-10, 10))+
  scale_x_continuous(limits = c(-10, 10))

g8

```

## SA- ASV raw data supp fig
```{r}

target_asvs <- res_filt_sa.join %>%
  filter(subdivision != "Dinoflagellata") %>%
  filter(is_significant_export & is_significant_transfer) %>%
  pull(ASV) 

ps_rel <- transform_sample_counts(deseq.sa.filt, function(x) x / sum(x)) %>%
  psmelt() 

plot_df <- ps_rel %>%
  filter(OTU %in% target_asvs)  %>%
  mutate(sample_depth = factor(sample_depth, 
                               levels = c("water_CTD", "trap_fix_1_2", "trap_fix_3_4"))) %>%
   mutate(label = paste0(subdivision, " | ", genus, "\n", OTU))

 p <- ggplot(plot_df, aes(x = sample_depth, y = Abundance, fill = sample_depth)) +
  geom_boxplot(alpha = 0.2, outlier.shape = NA) + 
  geom_jitter(width = 0.2, shape = 21, size = 3, alpha = 0.75) +
  facet_wrap(~label, scales = "free_y") + 
  theme_bw() +
  ggtitle("Subantarctic - Differentially abundant ASVs")+
  labs(y = "Relative Abundance", x = "Sample Depth")+
   theme(
    axis.text.x = element_text(angle = 45, hjust = 1), 
    legend.text = element_text(size = 12),
  legend.title = element_text(size = 14),
  legend.key.size = unit(1.5, "lines"))  
p
 # pdf("aldex_sa_raw rel abund.pdf", height= 9, width=17) ; plot(p) ; dev.off()
```


# Supplementrary figs
```{r}
p <- (g1 | g2) + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect')  + plot_annotation(
  title = 'Subtropical'
)
p
pdf("aldex_export transfer_ST.pdf", height= 7, width=10) ; plot(p) ; dev.off()

p <- (g5 | g6) + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect')  + plot_annotation(
  title = 'Subantarctic'
)
p
pdf("aldex_export transfer_SA.pdf", height= 7, width=10) ; plot(p) ; dev.off()

p <- (g1_dino | g2_dino) + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect')  + plot_annotation(
  title = 'Subtropical'
)
p
pdf("aldex_export transfer_ST_dino.pdf", height= 7, width=10) ; plot(p) ; dev.off()

p <- (g5_dino | g6_dino) + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect')  + plot_annotation(
  title = 'Subantarctic'
)
p
pdf("aldex_export transfer_SA_dino.pdf", height= 7, width=10) ; plot(p) ; dev.off()
```

# Main figure 7: deseq output
```{r}

g3 <- g3 + guides(fill = "none", colour = "none", shape = "none", size = "none")
g4 <- g4 + guides(fill = "none", colour = "none", shape = "none", size = "none")
g7 <- g7 + guides(fill = "none", colour = "none", shape = "none", size = "none")
g8 <- g8 + guides(fill = "none", colour = "none", shape = "none", size = "none")

p <- (g3 | g7) / (g4 | g8) + plot_annotation(tag_levels = 'A')
p
pdf("aldex_mainv2.pdf", height= 9, width=9) ; plot(p) ; dev.off()


```

# summary tables
```{r}

res_summary_st <- res_filt_st.join %>% 
  filter(is_significant_export & is_significant_transfer) %>%
  select(ASV, domain:species)

res_summary_sa <- res_filt_sa.join %>% 
  filter(is_significant_export & is_significant_transfer) %>%
  select(ASV, domain:species)

res_summary <- full_join(res_summary_st, res_summary_sa) %>%
  full_join(core_asvs %>% select(ASV, sediment)) %>%
  filter(!(is.na(division))) %>%
  distinct()

res_summary %>% group_by(subdivision,class) %>% summarise(count = n())

res_summary %>% group_by(sediment) %>% summarise(count = n()) 
```

## Number of common and uunique ASVs for water mass
```{r}
ST_asvs <- res_filt_st.join %>% 
  filter(is_significant_export & is_significant_transfer) %>%
  select(ASV) %>%
  mutate(water_mass = "ST")
SA_asvs <- res_filt_sa.join %>% 
  filter(is_significant_export & is_significant_transfer) %>%
  select(ASV) %>%
  mutate(water_mass = "SA")

wm_asvs <- full_join(ST_asvs, SA_asvs) 

# number of common and unique ASVs per water mass from deseq analysis
upset_df <- wm_asvs %>%
  mutate(n=1) %>%
  pivot_wider(names_from = water_mass,
              values_fill = 0,
              values_from = n) %>%
  data.frame() # this part is important at the end of forming the df if not will have error later

fig_upset_asv <- UpSetR::upset(upset_df, 
                                # nsets = 5,
                                # empty.intersections = "off", 
                                # nintersects = 23,
                                # number.angles = 30, 
                                # point.size = 10, 
                                # line.size = 4, 
                                mainbar.y.label = "ASV number", 
                                sets.x.label = "ASVs", 
                                order.by = "freq", 
                                # text.scale = c(3), 
                                mb.ratio = c(0.6, 0.4))
fig_upset_asv

```

## Supplementary table S3 - Number of ASVs by taxa, shared and common between water mass 
```{r}
taxa <- as.data.frame(ps@tax_table)
asv_df <- upset_df %>%
  mutate(group = case_when(ST == 1 & SA == 1 ~ "both",
                           ST == 1 & SA == 0 ~ "ST",
                           ST == 0 & SA == 1 ~ "SA")) %>%
  left_join(taxa)%>%
  full_join(core_asvs %>% select(ASV, sediment)) %>%
  filter(!(is.na(group))) %>%
  distinct() %>%
  mutate(sediment = coalesce(sediment, "not detected"))  # label sediment ASVs

# list of unique or shared ASVs between the groups
summary <- asv_df %>%
  # filter(ST == 1) %>%
  group_by(group,subdivision,class, order) %>%
  summarise(n=n()) %>%
  pivot_wider(names_from = group, values_from = n, values_fill = 0) %>%
  mutate(total = ST+SA+both) %>%
  arrange(subdivision, class, desc(total))
writexl::write_xlsx(summary, "aldex_taxa_summary.xlsx")

```

## number of ASVs detected in the sediments
```{r}
summary <- asv_df %>%
  group_by(sediment, subdivision) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  pivot_wider(names_from = "sediment", values_from = "n", values_fill = 0) %>%
  mutate(prop = detected / (detected + `not detected`)*100)

summary %>%
  summarise(
    detected_total = sum(detected, na.rm = TRUE),
    not_detected_total = sum(`not detected`, na.rm = TRUE),
    prop_total = detected_total / (detected_total + not_detected_total)
  )
```

