---
title: "deseq_v5"
author: "Denise Ong"
date: "2023-10-23"
output: html_document
---

Deseq2 code for export and transfer potential
For main figure 7 analysis
- compare CTD samples vs all fixed traps depth 1 and 2 vs all fixed traps 3 and 4
- separate water mass, but not interaction terms
- mark out ASVs that are detected in sediment samples
- include Dinoflagellata for analysis but present separately for main fig


# Initialize
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(here)

source(here("init_files", "init.R")) #for libraries
source(here("init_files", "init_markdown.R")) #for markdown libraries
source(here("init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps

set.seed(1)
library(EnhancedVolcano)

# Subdivision palette
divisionPalette <- readxl::read_excel("init_files/division_palette.xlsx")
divisionPalette <- structure(divisionPalette$colour,.Names=divisionPalette$subdivision) 
```


# load data
```{r}
source(here("init_files", "read_data_v3.R")) #load phyloseq and modify variables
ps_all
source(here("init_files", "filt_samples.R")) #filter out replicates, blanks etc. 
ps
```

# list of ASVs from sediment
```{r}
ps_core <- ps %>% # ready subsetted phyloseq file in microviz 2 Rmd, removing replicates, trap blanks and erroneous samples.
   ps_filter(sample == "core")
ps_norm <- phyloseq_normalize_median(ps_core)
ps_long <- phyloseq_transform_to_long(ps_norm)

core_asvs <- ps_long %>%
  select(ASV, water_mass) %>%
  unique() %>%
  mutate(sediment = "detected")

core_asvs_ST <- core_asvs %>%
  filter(water_mass == "ST")

core_asvs_SA <- core_asvs %>%
  filter(water_mass == "SA")
```


Notes for Deseq analysis:
dealing with 0 in OTU table (geometric means) - to use fittype = local
https://github.com/joey711/phyloseq/issues/642
https://github.com/joey711/phyloseq/issues/387
https://support.bioconductor.org/p/63229/

based on MA plot distribution looks similar, following recommendations of this page -> use filter out ASVs that do not occur in at least half of the samples. minRep=Inf and cooksCutoff=FALSE in combination with strict filtering and then use lfcShrink which adds a layer of stringency as well. check for outliers with PCA plot, inspect top genes with plotCounts.
https://support.bioconductor.org/p/9145823/ 
also here https://www.biostars.org/p/149031/
guide on how to filter out the ASVs: https://bioconductor.github.io/BiocWorkshops/rna-seq-data-analysis-with-deseq2.html

how to use lfcshrink- changes log2fold values to be more conservative, but does not change the p values.
https://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#extended-section-on-shrinkage-estimators
https://hbctraining.github.io/DGE_workshop/lessons/05_DGE_DESeq2_analysis2.html
23 Oct 2023: I am not using lfc shink anymore, it creates weird lfc values only for transfer part. After strict filtering, the values of LFC are similar to manually calculated ones.

http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#se
results table: the log2 fold change and Wald test p value will be for the last variable in the design formula

•	baseMean—The average of the normalized count values, dividing by size factors, taken over all samples.
•	log2FoldChange–The effect size estimate. This value indicates how much the gene or transcript's expression seems to have changed between the comparison and control groups. This value is reported on a logarithmic scale to base 2.
•	lfcSE–The standard error estimate for the log2 fold change estimate.
•	stat–The value of the test statistic for the gene or transcript.
•	pvalue–P-value of the test for the gene or transcript.
•	padj–Adjusted P-value for multiple testing for the gene or transcript.

- compare CTD samples vs all fixed traps depth 1 and 2 vs all fixed traps 3 and 4


# DESeq by water mass and sample type
create 6 groups instead
note: lfc shrink changes the lfc of transfer only to create weird values. Based on the manual calculation the normal LFC values is not that much different. Using the normal LFC values for this. 
## deseq analysis
```{r}
deseq.subset <- ps %>%
  # combine depth 1+2, 3+4
  ps_filter(sample == "water_CTD" | sample == "trap_fix") %>%
  ps_mutate(sample_depth = case_when(sample == "trap_fix" ~ paste(sample, depth_category, sep = "_"), 
                                     # add label for depth only for fixed traps
                                     sample == "water_CTD" ~ sample)) %>%
  ps_mutate(sample_depth = case_when(sample_depth == "trap_fix_1" ~ "trap_fix_1_2",
                                     sample_depth == "trap_fix_2" ~ "trap_fix_1_2",
                                     sample_depth == "trap_fix_3" ~ "trap_fix_3_4",
                                     sample_depth == "trap_fix_4" ~ "trap_fix_3_4",
                                     TRUE ~ as.character(sample_depth))) %>%
  ps_mutate(sample_depth_wm = paste(sample_depth, water_mass, sep = "_"))  %>%
  tax_filter(min_total_abundance = 11)

#no. of samples in each group
deseq.subset@sam_data %>%
  group_by(sample_depth_wm) %>%
  summarise(n=n())

diagdds = phyloseq_to_deseq2(deseq.subset, ~  sample_depth_wm)
diagdds$sample_depth_wm #check groups

# Minimal filtering to reduce the size of the dataset. https://bioconductor.github.io/BiocWorkshops/rna-seq-data-analysis-with-deseq2.html remove all ASVs that have a count of 10 or more for 4 or more samples (half of lowest no. of samples group trap_fix_3_4)
keep <- rowSums(counts(diagdds) >= 20) >= 4
table(keep)
diagdds <- diagdds[keep,]

# after filtering ASVs remove 1964 ASVs and keep 960 ASVs

#check number of reads in each sample
# boxplot(log10(counts(diagdds)+1))

# Estimate geometric mean with zeros - to deal with 0s in the ASV table
gm_mean = function(x, na.rm=TRUE){
    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
geoMeans = apply(counts(diagdds), 1, gm_mean)
diagdds = estimateSizeFactors(diagdds, geoMeans = geoMeans) 
# boxplot(log10(counts(diagdds,normalized=TRUE)+1))

# Run deseq function
diagdds = DESeq(diagdds, fitType="local", minReplicatesForReplace=Inf)

normalised_counts<- counts(diagdds, normalized=TRUE)
# plotDispEsts(diagdds) #  check dispersion plot. See https://hbctraining.github.io/DGE_workshop/lessons/04_DGE_DESeq2_analysis.html
# resultsNames(diagdds) # checks for the groups produced. Using results function, contrast variable to specify which groups to compare.

#PCA plot
# vsd <- vst(diagdds, blind=FALSE, nsub = 500)
# plotPCA(vsd, intgroup=c("sample_depth_wm"))
```
## ST - export efficiency
```{r}
res_export <- results(diagdds,
             contrast=c("sample_depth_wm", "trap_fix_1_2_ST", "water_CTD_ST"),
             cooksCutoff = FALSE
             )
plotMA(res_export, cex=.8)

EnhancedVolcano(res_export,
    lab = rownames(res_export),
    x = 'log2FoldChange',
    y = 'pvalue',
    pCutoff = 0.05,
    FCcutoff = 0.5)
# outliers > 10 and <-10 - removed after filtering for ASVs not present in any sample type
```

```{r}
# extract data frame, no filtering for p value
alpha = 1
sigtab_export = res_export[which(res_export$padj <= alpha), ]
sigtab_export = cbind(as(sigtab_export, "data.frame"), as(tax_table(ps)[rownames(sigtab_export), ], "matrix")) # change ps if ps input changes

#get mean counts for each ASV in each group
asv_means_ST <-data.frame()
for(i in sigtab_export$ASV){
  water<- mean(counts(diagdds, normalized=TRUE)[i,which(diagdds$sample_depth_wm == "water_CTD_ST")])
  upp_trap<- mean(counts(diagdds, normalized=TRUE)[i,which(diagdds$sample_depth_wm == "trap_fix_1_2_ST")])
  low_trap<- mean(counts(diagdds, normalized=TRUE)[i,which(diagdds$sample_depth_wm == "trap_fix_3_4_ST")])
  asv = i
  output = c(asv, water, upp_trap, low_trap)
  asv_means_ST = rbind(asv_means_ST, output)
  }
colnames(asv_means_ST) <- c("ASV", "CTD_means", "upp_trap_means", "low_trap_means")

#manual log2fold counts
asv_means_ST <- asv_means_ST %>%
  mutate(lcf_manual_export = log2(as.numeric(upp_trap_means)/as.numeric(CTD_means))) %>%
  mutate(lcf_manual_transfer = log2(as.numeric(low_trap_means)/as.numeric(upp_trap_means)))

sigtab_export_means_ST <- full_join(sigtab_export, asv_means_ST, by=join_by(ASV)) %>%
  unique()

# filter out ASVs not present in all sample types, and pump ASVs
sigtab_export_sub_ST <- sigtab_export_means_ST %>%
  filter(CTD_means != 0 & upp_trap_means != 0 & low_trap_means != 0) %>% # filter ASVs not present in any of the three sample types
  filter(!(is.na(log2FoldChange)))

# checks
# check with volcano plot
sigtab_export_sub_ST %>%
  ggplot(aes(x=log2FoldChange, y=-(log10(padj))))+
  geom_point()

#manual checks
# plotCounts(diagdds, gene = "asv_047_00084", intgroup="sample_depth_wm")
```

```{r}
sigtab_export_taxa <- sigtab_export_sub_ST %>%
  ### select only few taxa to plot at order level. Rest at class or family (Diatoms) level
  mutate(taxa = case_when(subdivision == "Dinoflagellata" ~ as.character(order),
                          # class == "Polycystinea" ~ as.character(order),
                          # class == "Acantharea" ~ as.character(order),
                          # class == "Prymnesiophyceae" ~ as.character(order),
                          # subdivision == "Gyrista" ~ as.character(order),
                          TRUE~ as.character(class))) 
sigtab_export_taxa$taxa <- as.character(sigtab_export_taxa$taxa)

# check <- sigtab_export_taxa %>%
#   group_by(taxa) %>%
#   summarise(n=n())

sigtab_export_taxa <- full_join(sigtab_export_taxa, core_asvs_ST) %>%
  mutate(sediment = coalesce(sediment, "not detected")) %>%
  filter(!(is.na(log2FoldChange))) 

sigtab_export_taxa <- sigtab_export_taxa %>%
  arrange(taxa)
#--------- Plot
# Division and Order plot. Need change this manually if want to
x = tapply(sigtab_export_taxa$log2FoldChange, sigtab_export_taxa$taxa, function(x) max(x))
x = sort(x, TRUE)
# sigtab_export_taxa$taxa = factor(as.character(sigtab_export_taxa$taxa), levels=names(x))
g1 <- sigtab_export_taxa %>%
  filter(subdivision != "Dinoflagellata") %>%
  ggplot(aes(x=fct_rev(fct_inorder(taxa)), y=log2FoldChange)) + 
  geom_point(aes(shape=sediment, size=sediment, fill=subdivision), colour="black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("log2 fold change - Water CTD <--> Upper fixed traps") +
  ggtitle("Export efficiency") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(subdivision~., scales = "free", space = "free", switch = "y") +
  # scale_shape_manual(values=c(3, 16, 17)) +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g1
# pdf("deseq_legend.pdf", height= 10, width=7) ; plot(g1) ; dev.off()

g1_dino <- sigtab_export_taxa %>%
  filter(subdivision == "Dinoflagellata") %>%
  ggplot(aes(x=fct_rev(fct_inorder(taxa)), y=log2FoldChange)) + 
  geom_point(aes(shape=sediment, size=sediment, fill=class), colour="black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("log2 fold change - Water CTD <--> Upper fixed traps") +
  ggtitle("Export efficiency") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(class~., scales = "free", space = "free", switch = "y") +
  # scale_shape_manual(values=c(3, 16, 17)) +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g1_dino

```


## ST - transfer efficiency

```{r}
res_transfer <- results(diagdds,
             contrast=c("sample_depth_wm", "trap_fix_3_4_ST", "trap_fix_1_2_ST"),
             cooksCutoff = FALSE)
plotMA(res_transfer, cex=.8)


# check volcano plots before filtering ASVs
EnhancedVolcano(res_transfer,
    lab = rownames(res_transfer),
    x = 'log2FoldChange',
    y = 'pvalue',
    pCutoff = 0.05,
    FCcutoff = 0.5)
```
```{r}
# extract data frame
alpha = 1
sigtab_transfer = res_transfer[which(res_transfer$padj <= alpha), ]
sigtab_transfer = cbind(as(sigtab_transfer, "data.frame"), as(tax_table(ps)[rownames(sigtab_transfer), ], "matrix")) # change ps if ps input changes

sigtab_transfer_means_ST <- dplyr::full_join(sigtab_transfer, asv_means_ST) %>%
  unique()

sigtab_transfer_sub_ST <- sigtab_transfer_means_ST %>%
  filter(CTD_means != 0 & upp_trap_means != 0 & low_trap_means !=0) %>% # filter ASVs not present in any of the three sample types
  filter(!(is.na(log2FoldChange)))

# check volcano plots after filtering ASVs
sigtab_transfer_sub_ST %>%
  ggplot(aes(x=log2FoldChange, y=-(log10(padj))))+
  geom_point()
  
```

```{r}

sigtab_transfer_taxa <- sigtab_transfer_sub_ST %>%
  ### select only few taxa to plot at order level. Rest at class or family (Diatoms) level
  mutate(taxa = case_when(subdivision == "Dinoflagellata" ~ as.character(order),
                          # class == "Polycystinea" ~ as.character(order),
                          # class == "Acantharea" ~ as.character(order),
                          # class == "Prymnesiophyceae" ~ as.character(order),
                          # subdivision == "Gyrista" ~ as.character(order),
                          TRUE~ as.character(class))) 
sigtab_transfer_taxa$taxa <- as.character(sigtab_transfer_taxa$taxa)
# check <- sigtab_transfer_taxa %>%
#   group_by(taxa) %>%
#   summarise(n=n())

sigtab_transfer_taxa <- full_join(sigtab_transfer_taxa, core_asvs_ST) %>%
  mutate(sediment = coalesce(sediment, "not detected")) %>%
  filter(!(is.na(log2FoldChange)))

sigtab_transfer_taxa <- sigtab_transfer_taxa %>%
  arrange(taxa)

# sigtab_transfer_taxa$taxa %>% levels()

#--------- Plot
# Division and Order plot. Need change this manually if want to
x = tapply(sigtab_transfer_taxa$log2FoldChange, sigtab_transfer_taxa$taxa, function(x) max(x))
x = sort(x, TRUE)
# sigtab_transfer_taxa$taxa = factor(as.character(sigtab_transfer_taxa$taxa), levels=names(x))
g2 <- sigtab_transfer_taxa %>%
  filter(subdivision != "Dinoflagellata") %>%
  ggplot(aes(x=fct_rev(fct_inorder(taxa)), y=log2FoldChange, fill=subdivision)) + 
  geom_point(aes(shape=sediment, size=sediment), colour = "black")+
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "none",
        strip.text.y = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(), 
        strip.background = element_blank()
        ) +
  ylab("log2 fold change - Upper fixed traps <--> Lower fixed traps") +
  ggtitle("Transfer efficiency") +
  facet_grid(subdivision~., scales = "free", space = "free", switch = "y") +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))

g2
# pdf("deseq_transfer_division_v2.pdf", height= 5, width=7) ; plot(g2) ; dev.off()


g2_dino <- sigtab_transfer_taxa %>%
  filter(subdivision == "Dinoflagellata") %>%
  ggplot(aes(x=fct_rev(fct_inorder(taxa)), y=log2FoldChange, fill=class)) + 
  geom_point(aes(shape=sediment, size=sediment), colour = "black")+
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))+
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 8),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "none",
        strip.text.y = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(), 
        strip.background = element_blank()
        ) +
  ylab("log2 fold change - Upper fixed traps <--> Lower fixed traps") +
  ggtitle("Transfer efficiency") +
  facet_grid(class~., scales = "free", space = "free", switch = "y") +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))

g2_dino
```
### Supplementrary figs
```{r}
p <- (g1 | g2) + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect')  + plot_annotation(
  title = 'Subtropical'
)
p
pdf("deseq_export transfer_ST_v7_no dino.pdf", height= 7, width=10) ; plot(p) ; dev.off()

p <- (g1_dino | g2_dino) + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect')  + plot_annotation(
  title = 'Subtropical'
)
p
pdf("deseq_export transfer_ST_v7_dino.pdf", height= 5, width=10) ; plot(p) ; dev.off()
```

## ST - Export vs transfer
```{r}
# export vs transfer efficiency plot. Only include ASVs that are significant in both DESEQ analysis.
sigtab_transfer <- sigtab_transfer_sub_ST %>%
  dplyr::rename("baseMean_trans" = "baseMean",
         "log2FoldChange_trans" = "log2FoldChange",
         "lfcSE_trans" = "lfcSE",
         "stat_trans" = "stat",
         "pvalue_trans" = "pvalue",
         "padj_trans" = "padj")
sigtab_export <- sigtab_export_sub_ST %>%
  dplyr::rename("baseMean_exp" = "baseMean",
         "log2FoldChange_exp" = "log2FoldChange",
         "lfcSE_exp" = "lfcSE",
         "stat_exp" = "stat",
         "pvalue_exp" = "pvalue",
         "padj_exp" = "padj")

sigtab_export_transfer_ST <- full_join(sigtab_transfer, sigtab_export) %>% # join export and transfer tables
  drop_na() %>%
  full_join(core_asvs_ST) %>%
  mutate(sediment = coalesce(sediment, "not detected")) %>% # add sediment ASVs labels
  filter(!(is.na(log2FoldChange_exp))) %>%
  ### select only few taxa to plot at order level. Rest at class or family (Diatoms) level
  mutate(taxa = case_when(subdivision == "Dinoflagellata" ~ as.character(order),
                          # class == "Polycystinea" ~ as.character(order),
                          # class == "Acantharea" ~ as.character(order),
                          # class == "Prymnesiophyceae" ~ as.character(order),
                          # subdivision == "Gyrista" ~ as.character(order),
                          TRUE~ as.character(class))) 

# writexl::write_xlsx(sigtab_export_transfer_ST, path = "deseq_sigtab_export_transfer_ST.xlsx")

# sub <- sigtab_export_transfer_ST %>%
#   filter(log2FoldChange_exp >= -5 & log2FoldChange_exp <= 5) %>%
#   filter(log2FoldChange_trans >= -5 & log2FoldChange_trans <= 5) 

sigtab_export_transfer_ST_sub <- sigtab_export_transfer_ST %>%
  filter(subdivision != "Dinoflagellata") 

g3 <- sigtab_export_transfer_ST_sub %>%
  ggplot(aes(x = log2FoldChange_exp, y = log2FoldChange_trans, label=class)) +
  geom_point(aes(shape=sediment, size=sediment, fill = subdivision), colour = "black")+
    geom_text_repel(
    data = subset(sigtab_export_transfer_ST_sub, sediment == "detected"),
    aes(label = class, colour = subdivision),
    size = 2.5, max.overlaps = 30,
    box.padding = 0.25,
    point.padding = 0.3,
    hjust = "outward",
    direction = "y"
    # size=2, max.overlaps = 10
                  )+
  geom_vline(xintercept = 0, linetype="solid") +  
  geom_hline(yintercept = 0, linetype="solid") +  
  # ggtitle("Export vs transfer efficiency") +
  xlab("Export efficiency log2 fold change - Water CTD <--> Upper fixed traps") +
  ylab("Transfer efficiency log2 fold change - Upper <--> Lower fixed traps") +
  scale_fill_manual(values=divisionPalette)+
  scale_colour_manual(values=divisionPalette)+
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        # legend.position = "bottom"
        legend.position = "none") +
  scale_size_manual(values=c(2.5, 2)) +
  scale_shape_manual(values = c(24,21))+
  coord_fixed() +
  ggtitle("Subtropical") +
  scale_y_continuous(limits = c(-10, 10))+
  scale_x_continuous(limits = c(-10, 10))

g3

sigtab_export_transfer_ST_dino<- sigtab_export_transfer_ST %>%
  filter(subdivision == "Dinoflagellata")

g3_dino <- sigtab_export_transfer_ST_dino %>%
  filter(subdivision == "Dinoflagellata") %>%
  ggplot(aes(x = log2FoldChange_exp, y = log2FoldChange_trans, label=class)) +
  geom_point(aes(shape=sediment, size=sediment, fill = class), colour = "black")+
    geom_text_repel(
    data = subset(sigtab_export_transfer_ST_dino, sediment == "detected"),
    aes(label = taxa, colour = class),
    size = 2.5, max.overlaps = 30,
    box.padding = 0.25,
    point.padding = 0.3,
    hjust = "outward",
    direction = "y")+
  geom_vline(xintercept = 0, linetype="solid") +  
  geom_hline(yintercept = 0, linetype="solid") +  
  # ggtitle("Export vs transfer efficiency") +
  xlab("Export efficiency log2 fold change - Water CTD <--> Upper fixed traps") +
  ylab("Transfer efficiency log2 fold change - Upper <--> Lower fixed traps") +
  scale_fill_manual(values=divisionPalette)+
  scale_colour_manual(values=divisionPalette)+
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        # legend.position = "bottom"
        legend.position = "none") +
  scale_size_manual(values=c(2.5, 2)) +
  scale_shape_manual(values = c(24,21))+
  coord_fixed() +
  ggtitle("Subtropical") +
  scale_y_continuous(limits = c(-10, 10))+
  scale_x_continuous(limits = c(-10, 10))

g3_dino
```

## SA - transfer efficiency

```{r}
res_transfer <- results(diagdds,
             contrast=c("sample_depth_wm", "trap_fix_3_4_SA", "trap_fix_1_2_SA"),
             cooksCutoff = FALSE)
plotMA(res_transfer, cex=.8)
# check volcano plots before filtering ASVs
EnhancedVolcano(res_transfer,
    lab = rownames(res_transfer),
    x = 'log2FoldChange',
    y = 'pvalue',
    pCutoff = 0.05,
    FCcutoff = 0.5)
```
```{r}
# extract data frame
alpha = 1
sigtab_transfer = res_transfer[which(res_transfer$padj <= alpha), ]
sigtab_transfer = cbind(as(sigtab_transfer, "data.frame"), as(tax_table(ps)[rownames(sigtab_transfer), ], "matrix")) # change ps if ps input changes

# put transfer first because there are some ASVs in transfer not detected in export
#get mean counts for each ASV in each group
asv_means_SA <-data.frame()
for(i in sigtab_transfer$ASV){
  water<- mean(counts(diagdds, normalized=TRUE)[i,which(diagdds$sample_depth_wm == "water_CTD_SA")])
  upp_trap<- mean(counts(diagdds, normalized=TRUE)[i,which(diagdds$sample_depth_wm == "trap_fix_1_2_SA")])
  low_trap<- mean(counts(diagdds, normalized=TRUE)[i,which(diagdds$sample_depth_wm == "trap_fix_3_4_SA")])
  asv = i
  output = c(asv, water, upp_trap, low_trap)
  asv_means_SA = rbind(asv_means_SA, output)
  }
colnames(asv_means_SA) <- c("ASV", "CTD_means", "upp_trap_means", "low_trap_means")

#manual log2fold counts
asv_means_SA <- asv_means_SA %>%
  mutate(lcf_manual_export = log2(as.numeric(upp_trap_means)/as.numeric(CTD_means))) %>%
  mutate(lcf_manual_transfer = log2(as.numeric(low_trap_means)/as.numeric(upp_trap_means)))


sigtab_transfer_means_SA <- full_join(sigtab_transfer, asv_means_SA)

sigtab_transfer_sub_SA <- sigtab_transfer_means_SA %>%
  filter(CTD_means != 0 & upp_trap_means != 0 & low_trap_means !=0)

# check volcano plots after filtering ASVs
sigtab_transfer_sub_SA %>%
  ggplot(aes(x=log2FoldChange, y=-(log10(padj))))+
  geom_point()
  
```

```{r}
sigtab_transfer_taxa <- sigtab_transfer_sub_SA %>%
  ### select only few taxa to plot at order level. Rest at class or family (Diatoms) level
  mutate(taxa = case_when(subdivision == "Dinoflagellata" ~ as.character(order),
                          # class == "Prymnesiophyceae" ~ as.character(order),
                          # class == "Polycystinea" ~ as.character(order),
                          # class == "Acantharea" ~ as.character(order),
                          # subdivision == "Gyrista" ~ as.character(order),
                          TRUE~ as.character(class))) 

sigtab_transfer_taxa$taxa <- as.character(sigtab_transfer_taxa$taxa)

sigtab_transfer_taxa <- full_join(sigtab_transfer_taxa, core_asvs_SA) %>%
  mutate(sediment = coalesce(sediment, "not detected")) %>%
  filter(!(is.na(log2FoldChange)))

sigtab_transfer_taxa <- sigtab_transfer_taxa %>%
  arrange(taxa)
# sigtab_transfer_taxa <- sigtab_transfer_taxa %>%
#   filter(log2FoldChange >= -5)
#--------- Plot
# Division and Order plot. Need change this manually if want to
x = tapply(sigtab_transfer_taxa$log2FoldChange, sigtab_transfer_taxa$taxa, function(x) max(x))
x = sort(x, TRUE)
# sigtab_transfer_taxa$taxa = factor(as.character(sigtab_transfer_taxa$taxa), levels=names(x))
g5 <- sigtab_transfer_taxa %>%
  filter(subdivision != "Dinoflagellata") %>%
  ggplot(aes(x=fct_rev(fct_inorder(taxa)), y=log2FoldChange, fill=subdivision)) + 
  geom_point(aes(shape=sediment, size=sediment)) +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) +
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 7),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("log2 fold change - Upper fixed traps <--> Lower fixed traps") +
  ggtitle("Transfer efficiency") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(subdivision~., scales = "free", space = "free", switch = "y") +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24, 21))
g5

# pdf("deseq_transfer_division_v2.pdf", height= 5, width=7) ; plot(g2) ; dev.off()

g5_dino <- sigtab_transfer_taxa %>%
  filter(subdivision == "Dinoflagellata") %>%
  ggplot(aes(x=fct_rev(fct_inorder(taxa)), y=log2FoldChange, fill=class)) + 
  geom_point(aes(shape=sediment, size=sediment)) +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) +
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 7),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("log2 fold change - Upper fixed traps <--> Lower fixed traps") +
  ggtitle("Transfer efficiency") +
  theme(strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(class~., scales = "free", space = "free", switch = "y") +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24, 21))
g5_dino

```

## SA - export efficiency
```{r}
res_export <- results(diagdds,
                      contrast=c("sample_depth_wm", "trap_fix_1_2_SA", "water_CTD_SA"),
                      cooksCutoff = FALSE
                      )
plotMA(res_export, cex=.8)
EnhancedVolcano(res_export,
    lab = rownames(res_export),
    x = 'log2FoldChange',
    y = 'pvalue',
    pCutoff = 0.05,
    FCcutoff = 0.5)
```

```{r}
# extract data frame
alpha = 1
sigtab_export = res_export[which(res_export$padj <= alpha), ]
sigtab_export = cbind(as(sigtab_export, "data.frame"), as(tax_table(ps)[rownames(sigtab_export), ], "matrix")) # change ps if ps input changes

sigtab_export_means_SA <- full_join(sigtab_export, asv_means_SA)

sigtab_export_sub_SA <- sigtab_export_means_SA %>%
  filter(CTD_means != 0 & upp_trap_means != 0 & low_trap_means != 0)

sigtab_export_sub_SA %>%
  ggplot(aes(x=log2FoldChange, y=-(log10(padj))))+
  geom_point()
```

```{r}
sigtab_export_taxa <- sigtab_export_sub_SA %>%
  ### select only few taxa to plot at order level. Rest at class or family (Diatoms) level
  mutate(taxa = case_when(subdivision == "Dinoflagellata" ~ as.character(order),
                          # class == "Polycystinea" ~ as.character(order),
                          # class == "Acantharea" ~ as.character(order),
                          # class == "Prymnesiophyceae" ~ as.character(order),
                          # subdivision == "Gyrista" ~ as.character(order),
                          TRUE~ as.character(class))) 
sigtab_transfer_taxa$taxa <- as.character(sigtab_transfer_taxa$taxa)

sigtab_export_taxa <- full_join(sigtab_export_taxa, core_asvs_SA) %>%
  mutate(sediment = coalesce(sediment, "not detected")) %>%
  filter(!(is.na(log2FoldChange)))

sigtab_export_taxa <- sigtab_export_taxa %>%
  arrange(taxa)

#--------- Plot
# Division and Order plot. Need change this manually if want to
x = tapply(sigtab_export_taxa$log2FoldChange, sigtab_export_taxa$taxa, function(x) max(x))
x = sort(x, TRUE)
# sigtab_export_taxa$taxa = factor(as.character(sigtab_export_taxa$taxa), levels=names(x))
g4 <- sigtab_export_taxa %>%
  filter(subdivision != "Dinoflagellata") %>%
  ggplot(aes(x=fct_rev(fct_inorder(taxa)), y=log2FoldChange, fill=subdivision)) + 
  geom_point(aes(shape=sediment, size=sediment), colour = "black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) +
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 7),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("log2 fold change - Water CTD <--> Upper fixed traps") +
  ggtitle("Export efficiency") +
  theme(
        # axis.text.x = element_text(angle=90),
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(subdivision~., scales = "free", space = "free", switch = "y") +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g4
# pdf("deseq_legend.pdf", height= 10, width=7) ; plot(g1) ; dev.off()

g4_dino <- sigtab_export_taxa %>%
  filter(subdivision == "Dinoflagellata") %>%
  ggplot(aes(x=fct_rev(fct_inorder(taxa)), y=log2FoldChange, fill=class)) + 
  geom_point(aes(shape=sediment, size=sediment), colour = "black") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5)) +
  scale_fill_manual(values=divisionPalette) +
  geom_hline(yintercept = 0, linetype="solid") +
  coord_flip() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 8),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 7),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right"
        # legend.position = "none"
        ) +
  ylab("log2 fold change - Water CTD <--> Upper fixed traps") +
  ggtitle("Export efficiency") +
  theme(
        # axis.text.x = element_text(angle=90),
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.x = unit(0,"lines"),
        panel.spacing.y = unit(0, "lines"),
        panel.border = element_blank(),
        legend.title = element_markdown()) +
  facet_grid(class~., scales = "free", space = "free", switch = "y") +
  scale_size_manual(values=c(3, 2.5)) +
  scale_shape_manual(values = c(24,21))
g4_dino

```



### Supp figs
```{r}
p <- (g4 | g5) + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect') + plot_annotation(
  title = 'Subantarctic'
)
p
pdf("deseq_export transfer_SA_v7_no dino.pdf", height= 7, width=10) ; plot(p) ; dev.off()

p <- (g4_dino | g5_dino) + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect') + plot_annotation(
  title = 'Subantarctic'
)
p
pdf("deseq_export transfer_SA_v7_dino.pdf", height= 5, width=10) ; plot(p) ; dev.off()
```

## SA - Export vs transfer
```{r}
# export vs transfer efficiency plot. Only include ASVs that are significant in both DESEQ analysis.
sigtab_transfer <- sigtab_transfer_sub_SA %>%
  dplyr::rename("baseMean_trans" = "baseMean",
         "log2FoldChange_trans" = "log2FoldChange",
         "lfcSE_trans" = "lfcSE",
         "stat_trans" = "stat",
         "pvalue_trans" = "pvalue",
         "padj_trans" = "padj")
sigtab_export <- sigtab_export_sub_SA %>%
  dplyr::rename("baseMean_exp" = "baseMean",
         "log2FoldChange_exp" = "log2FoldChange",
         "lfcSE_exp" = "lfcSE",
         "stat_exp" = "stat",
         "pvalue_exp" = "pvalue",
         "padj_exp" = "padj")
sigtab_export_transfer_SA <- full_join(sigtab_transfer, sigtab_export) %>%
  drop_na() %>%
  full_join(core_asvs_SA) %>%
  mutate(sediment = coalesce(sediment, "not detected")) %>% # label sediment ASVs
  filter(!(is.na(log2FoldChange_trans)))%>%
  ### select only few taxa to plot at order level. Rest at class or family (Diatoms) level
  mutate(taxa = case_when(subdivision == "Dinoflagellata" ~ as.character(order),
                          # class == "Polycystinea" ~ as.character(order),
                          # class == "Acantharea" ~ as.character(order),
                          # class == "Prymnesiophyceae" ~ as.character(order),
                          # subdivision == "Gyrista" ~ as.character(order),
                          TRUE~ as.character(class)))

# writexl::write_xlsx(sigtab_export_transfer_SA, path = "deseq_sigtab_export_transfer_SA.xlsx")

sigtab_export_transfer_SA_sub <- sigtab_export_transfer_SA %>%
  filter(subdivision != "Dinoflagellata")

g6 <- sigtab_export_transfer_SA_sub %>%
  ggplot(aes(x = log2FoldChange_exp, y = log2FoldChange_trans, label=class)) +
  geom_point(aes(shape=sediment, size=sediment, fill = subdivision), colour = "black")+
  geom_text_repel(
    data = subset(sigtab_export_transfer_SA_sub, sediment == "detected"),
    aes(label = taxa, colour = subdivision),
    size = 2.5, max.overlaps = 30,
    box.padding = 0.5,
    point.padding = 0,
    hjust = "outward",
    direction = "y"
    # min.segment.length = 0.25
    # size=2, max.overlaps = 10
                  )+
  geom_vline(xintercept = 0, linetype="solid") +  
  geom_hline(yintercept = 0, linetype="solid") +  
  # ggtitle("Export vs transfer efficiency") +
  xlab("Export efficiency log2 fold change - Water CTD <--> Upper fixed traps") +
  ylab("Transfer efficiency log2 fold change - Upper <--> Lower fixed traps") +
  scale_colour_manual(values=divisionPalette)+
  scale_fill_manual(values=divisionPalette)+
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right") +
  coord_fixed() +
  ggtitle("Subantarctic")+
  scale_size_manual(values=c(2.5,2)) +
  scale_shape_manual(values = c(24,21)) +
  scale_y_continuous(limits = c(-10, 10))+
  scale_x_continuous(limits = c(-10, 10))

g6

sigtab_export_transfer_SA_dino <- sigtab_export_transfer_SA %>%
  filter(subdivision == "Dinoflagellata")

g6_dino <- sigtab_export_transfer_SA_dino %>%  
  ggplot(aes(x = log2FoldChange_exp, y = log2FoldChange_trans, label=class)) +
  geom_point(aes(shape=sediment, size=sediment, fill = class), colour = "black")+
  geom_text_repel(
    data = subset(sigtab_export_transfer_SA_dino, sediment == "detected"),
    aes(label = taxa, colour = class),
    size = 2.5, max.overlaps = 30,
    box.padding = 0.5,
    point.padding = 0,
    hjust = "outward",
    direction = "y"
    # min.segment.length = 0.25
    # size=2, max.overlaps = 10
                  )+
  geom_vline(xintercept = 0, linetype="solid") +  
  geom_hline(yintercept = 0, linetype="solid") +  
  # ggtitle("Export vs transfer efficiency") +
  xlab("Export efficiency log2 fold change - Water CTD <--> Upper fixed traps") +
  ylab("Transfer efficiency log2 fold change - Upper <--> Lower fixed traps") +
  scale_colour_manual(values=divisionPalette)+
  scale_fill_manual(values=divisionPalette)+
  theme(axis.text.x = element_text(size = 10),
        axis.title.x =element_text(size=10),
        axis.text.y = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.position = "right") +
  coord_fixed() +
  ggtitle("Subantarctic")+
  scale_size_manual(values=c(2.5,2)) +
  scale_shape_manual(values = c(24,21)) +
  scale_y_continuous(limits = c(-10, 10))+
  scale_x_continuous(limits = c(-10, 10))

g6_dino
```

# Main figure 7: deseq output
```{r}

# p <- g3 | g6 + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect')
# p
# pdf("deseq_combine_v7_no dino.pdf", height= 6, width=15) ; plot(p) ; dev.off()
# 
# p <- g3_dino | g6_dino + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect')
# p
# pdf("deseq_combine_v7_dino.pdf", height= 6, width=15) ; plot(p) ; dev.off()

p <- ((g3 | g6) / (g3_dino | g6_dino)) + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect')
p
pdf("deseq_combine_v7_all.pdf", height= 9, width=11) ; plot(p) ; dev.off()


```

# Supplementary table S4: Summary table for deseq LFC output
```{r}
summary_ST <- sigtab_export_transfer_ST %>%
  mutate(subdivision = case_when(subdivision == "Dinoflagellata" ~ as.character(class),
                                 subdivision == "Chlorophyta_X" ~ as.character(class),
                                 subdivision == "Haptophyta_X" ~ as.character(order),
                          TRUE~ as.character(subdivision))) %>%
  select(subdivision,log2FoldChange_exp,log2FoldChange_trans) %>%
  filter(!(subdivision %in% c("Cercozoa","Picozoa_X", "Cryptophyta_X", "Noctilucophyceae"))) %>%
  group_by(subdivision) %>%
  summarise(mean_exp_ST = round(mean(log2FoldChange_exp),2),
            sd_exp_ST = round(sd(log2FoldChange_exp),2),
            mean_trans_ST = round(mean(log2FoldChange_trans),2),
            sd_trans_ST = round(sd(log2FoldChange_trans),2),
            n_ST=n()) %>%
  unite("exp_ST", mean_exp_ST:sd_exp_ST, sep = " ± ") %>%
  unite("trans_ST", mean_trans_ST:sd_trans_ST, sep = " ± ")

summary_SA <- sigtab_export_transfer_SA %>%
  mutate(subdivision = case_when(subdivision == "Dinoflagellata" ~ as.character(class),
                                 subdivision == "Chlorophyta_X" ~ as.character(class),
                                 subdivision == "Haptophyta_X" ~ as.character(order),
                          TRUE~ as.character(subdivision))) %>%
  select(subdivision,log2FoldChange_exp,log2FoldChange_trans) %>%
  filter(!(subdivision %in% c("Cercozoa","Picozoa_X", "Cryptophyta_X", "Noctilucophyceae", "Bigyra", "Opisthokonta_X", "Choanoflagellata"))) %>%
  group_by(subdivision) %>%
  summarise(mean_exp_SA = round(mean(log2FoldChange_exp),2),
            sd_exp_SA = round(sd(log2FoldChange_exp),2),
            mean_trans_SA = round(mean(log2FoldChange_trans),2),
            sd_trans_SA = round(sd(log2FoldChange_trans),2),
            n_SA=n()) %>%
  unite("exp_SA", mean_exp_SA:sd_exp_SA, sep = " ± ") %>%
  unite("trans_SA", mean_trans_SA:sd_trans_SA, sep = " ± ")

summary <- full_join(summary_ST,summary_SA)

# writexl::write_xlsx(summary, path = "deseq_meanLFC.xlsx")
```

# Checks


## distribution of ASVs per quadrant
```{r}
# summary export transfer ratio
summary_ST <- sigtab_export_transfer_ST %>%
  mutate(export = case_when(log2FoldChange_exp >= 0 ~ "high",
                            log2FoldChange_exp < 0 ~ "low")) %>%
  mutate(transfer = case_when(log2FoldChange_trans >= 0 ~ "high",
                          log2FoldChange_trans < 0 ~ "low")) %>%
  group_by(class, export, transfer) %>%
  summarise(n=n())

summary_SA <- sigtab_export_transfer_SA %>%
  mutate(export = case_when(log2FoldChange_exp >= 0 ~ "high",
                            log2FoldChange_exp < 0 ~ "low")) %>%
  mutate(transfer = case_when(log2FoldChange_trans >= 0 ~ "high",
                          log2FoldChange_trans < 0 ~ "low")) %>%
  group_by(class, export, transfer) %>%
  summarise(n=n())

```


## Number of common and uunique ASVs for water mass
```{r}
ST_asvs <- sigtab_export_transfer_ST %>%
  select(ASV) %>%
  mutate(water_mass = "ST")
SA_asvs <- sigtab_export_transfer_SA %>%
  select(ASV) %>%
  mutate(water_mass = "SA")

wm_asvs <- full_join(ST_asvs, SA_asvs) 

# number of common and unique ASVs per water mass from deseq analysis
upset_df <- wm_asvs %>%
  mutate(n=1) %>%
  pivot_wider(names_from = water_mass,
              values_fill = 0,
              values_from = n) %>%
  data.frame() # this part is important at the end of forming the df if not will have error later

fig_upset_asv <- UpSetR::upset(upset_df, 
                                # nsets = 5,
                                # empty.intersections = "off", 
                                # nintersects = 23,
                                # number.angles = 30, 
                                # point.size = 10, 
                                # line.size = 4, 
                                mainbar.y.label = "ASV number", 
                                sets.x.label = "ASVs", 
                                order.by = "freq", 
                                # text.scale = c(3), 
                                mb.ratio = c(0.6, 0.4))
fig_upset_asv

```

## Supplementary table S3 - Number of ASVs by taxa, shared and common between water mass 
```{r}
taxa <- as.data.frame(ps@tax_table)
asv_df <- upset_df %>%
  mutate(group = case_when(ST == 1 & SA == 1 ~ "both",
                           ST == 1 & SA == 0 ~ "ST",
                           ST == 0 & SA == 1 ~ "SA")) %>%
  left_join(taxa)%>%
  ### select only few taxa to plot at order level. Rest at class or family (Diatoms) level
  mutate(taxa = case_when(class == "Dinophyceae" ~ as.character(order),
                          class == "Syndiniales" ~ as.character(order),
                          class == "Polycystinea" ~ as.character(order),
                          class == "Acantharea" ~ as.character(order),
                          class == "Prymnesiophyceae" ~ as.character(order),
                          subdivision == "Gyrista" ~ as.character(order),
                          TRUE~ as.character(class))) %>%
  full_join(core_asvs) %>%
  filter(!(is.na(group))) %>%
  mutate(sediment = coalesce(sediment, "not detected")) %>% # label sediment ASVs
  mutate(taxa = case_when(sediment == "detected" ~ taxa,
                          sediment == "not detected" ~ NA)) %>%
  select(-water_mass) %>%
  unique()

# list of unique or shared ASVs between the groups
summary <- asv_df %>%
  # filter(ST == 1) %>%
  group_by(group,subdivision,class) %>%
  summarise(n=n()) %>%
  pivot_wider(names_from = group, values_from = n, values_fill = 0) %>%
  mutate(total = ST+SA+both) %>%
  arrange(subdivision, desc(total))
# writexl::write_xlsx(summary, "deseq_taxa_summary.xlsx")

```

## number of ASVs detected in the sediments
```{r}
summary <- asv_df %>%
  # filter(ST == 1) %>%
  group_by(sediment, subdivision) %>%
  summarise(n=n()) 
```

## changes in rel abund for each taxa across the depths
```{r}
library(ggridges)

test_df <- sigtab_export_transfer_ST %>%
  mutate(subdivision = case_when(subdivision == "Dinoflagellata" ~ as.character(class),
                                 class == "Chloropicophyceae" ~ "Chloropicophyceae",
                                 order == "Isochrysidales" ~ "Isochrysidales",
                          TRUE~ as.character(subdivision))) %>%
  select(CTD_means,upp_trap_means,low_trap_means,subdivision) %>%
  mutate(CTD_means = as.numeric(CTD_means)) %>%
  mutate(upp_trap_means = as.numeric(upp_trap_means)) %>%
  mutate(low_trap_means = as.numeric(low_trap_means)) %>%
  filter(!(subdivision %in% c("Cercozoa","Picozoa_X", "Cryptophyta_X"))) %>%
  group_by(subdivision) %>%
  summarise_at(c("CTD_means", "upp_trap_means", "low_trap_means"), mean, na.rm = TRUE) %>%
  pivot_longer(cols=CTD_means:low_trap_means,
               names_to = "depth",
               values_to = "rel_abund") %>%
  group_by(subdivision) %>%
  mutate(density = rel_abund / sum(rel_abund)) %>%
  mutate(depth_num = case_when(depth == "CTD_means" ~3,
                               depth == "upp_trap_means" ~2,
                               depth == "low_trap_means" ~1)) 

p1 <- test_df %>%
  ggplot(aes(x = depth_num, y = fct_relevel(subdivision,
                                   "Chlorophyta_X", 
                                   "Chloropicophyceae",
                                   "Prasinodermophyta_X",	
                                   "Gyrista",
                                   "Haptophyta_X",
                                   "Isochrysidales",
                                   "Radiolaria",	
                                   "Ciliophora"), height = density, fill = subdivision)) + 
  geom_density_ridges(scale = 1, stat = "identity", alpha = 0.9)+
  coord_flip() +
  # facet_grid(~subdivision, scales = "free") +
  scale_colour_manual(values=divisionPalette)+
  scale_fill_manual(values=divisionPalette) +
  ylab("Subdivision") +
  ggtitle("Subtropical")+
  theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1),
        legend.position = "none")
  

p1

test_df <- sigtab_export_transfer_SA %>%
  mutate(subdivision = case_when(subdivision == "Dinoflagellata" ~ as.character(class),
                                 class == "Chloropicophyceae" ~ "Chloropicophyceae",
                                 order == "Isochrysidales" ~ "Isochrysidales",
                          TRUE~ as.character(subdivision))) %>%
  select(CTD_means,upp_trap_means,low_trap_means,subdivision) %>%
  mutate(CTD_means = as.numeric(CTD_means)) %>%
  mutate(upp_trap_means = as.numeric(upp_trap_means)) %>%
  mutate(low_trap_means = as.numeric(low_trap_means)) %>%
  group_by(subdivision) %>%
  # summarise(across(everything(), list(mean = mean)))
  summarise_at(c("CTD_means", "upp_trap_means", "low_trap_means"), mean, na.rm = TRUE) %>%
  pivot_longer(cols=CTD_means:low_trap_means,
               names_to = "depth",
               values_to = "rel_abund") %>%
  group_by(subdivision) %>%
  mutate(density = rel_abund / sum(rel_abund)) %>%
  mutate(depth_num = case_when(depth == "CTD_means" ~3,
                               depth == "upp_trap_means" ~2,
                               depth == "low_trap_means" ~1)) %>%
  filter(!(subdivision %in% c("Bigyra","Cercozoa", "Choanoflagellata", "Opisthokonta_X", "Picozoa_X", "Cryptophyta_X")))

p2 <- test_df %>%
  ggplot(aes(x = depth_num, y = fct_relevel(subdivision,
                                   "Chlorophyta_X", 	
                                   "Chloropicophyceae",
                                   "Prasinodermophyta_X",	
                                   "Gyrista",
                                   "Haptophyta_X",
                                   "Isochrysidales",
                                   "Radiolaria",	
                                   "Ciliophora"), height = density, fill = subdivision)) + 
  geom_density_ridges(scale = 1,stat = "identity", alpha = 0.9)+ 
  coord_flip() +
  # facet_grid(~subdivision, scales = "free") +
  scale_fill_manual(values=divisionPalette) +
  ylab("Subdivision")+
  ggtitle("Subantarctic")+
  theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1),
        legend.position = "none")


p2

p <- p1|p2
p
# pdf("deseq_export transfer_relabund.pdf", height=5, width=12) ; plot(p) ; dev.off()

```

